### 2013-12-13 manuscript.tex
added ignore file, added first scheleton of paper

% Created 2013-12-13 Fri 11:53
\documentclass[]{article}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{fixltx2e}
\usepackage{graphicx}
\usepackage{longtable}
\usepackage{float}
\usepackage{wrapfig}
\usepackage{rotating}
\usepackage[normalem]{ulem}
\usepackage{amsmath}
\usepackage{textcomp}
\usepackage{marvosym}
\usepackage{wasysym}
\usepackage{amssymb}
\usepackage{hyperref}
\tolerance=1000
\author{Daniele Barchiesi}
\date{}
\title{Dictionary learning for classification: learning incoherent subspaces}
\hypersetup{
  pdfkeywords={},
  pdfsubject={},
  pdfcreator={Emacs 24.3.1 (Org mode 8.2.4)}}
\begin{document}

\maketitle

\section{Abstract}
\label{sec-1}
\section{Introduction}
\label{sec-2}
\section{Background}
\label{sec-3}
\section{Incoherent Dictionary Learning}
\label{sec-4}
\section{Learning Incoherent Subspaces}
\label{sec-5}
\section{Numerical Experiments}
\label{sec-6}
\section{Discussion}
\label{sec-7}
\section{Conclusions}
\label{sec-8}
% Emacs 24.3.1 (Org mode 8.2.4)
\end{document}

### 2013-12-13 manuscript.tex
added initial document structure

@@
-% Created 2013-12-13 Fri 11:53
+% Created 2013-12-13 Fri 14:51
 \documentclass[]{article}
@@
 \author{Daniele Barchiesi}
-\date{}
-\title{Dictionary learning for classification: learning incoherent subspaces}
+\date{\textit{<2013-12-13 Fri>}}
+\title{Classification via dictionary learning: incoherent sub-space modelling for high-dimensional data}
 \hypersetup{
@@
 \maketitle
+\begin{abstract}
 
-\section{Abstract}
+\end{abstract}
+
+\section{Introduction}
 \label{sec-1}
-\section{Introduction}
+\begin{enumerate}
+\item $\square$ prepare a reading list
+\item $\square$ read papers
+\item $\square$ select relevant publications
+\item $\square$ write introduction and summaries
+\end{enumerate}
+\subsection{Problem definition}
+\label{sec-1-1}
+
+\subsection{Previous work and applications}
+\label{sec-1-2}
+
+\subsection{Outline and main contributions}
+\label{sec-1-3}
+
+\section{Background}
 \label{sec-2}
-\section{Background}
+\subsection{Benchmark techniques}
+\label{sec-2-1}
+\subsubsection{Features transform for classification}
+\label{sec-2-1-1}
+Feature transforms have been used to model data in an attempt to enhance the tradeoff between generalisation and discrimination, as described in Section \ref{sec-1}.
+Two of the main feature transform techniques include principal component analysis (PCA) and Fisher's linear discriminant analysis (LDA). This section provides a brief description of their rationale.
+
+Let $\{x_m \in R^N\}$ be a set of vectors containing features extracted from $M$ training signals.
+
+\subsubsection{Manifold learning}
+\label{sec-2-1-2}
+
+\subsection{Dictionary learning}
+\label{sec-2-2}
+\subsubsection{Dictionary learning for sparse approximation}
+\label{sec-2-2-1}
+\subsubsection{Incoherent dictionary learning}
+\label{sec-2-2-2}
+\subsubsection{Dictionary learning for classification}
+\label{sec-2-2-3}
+
+\section{Learning incoherent subspaces using iterative projections and rotations}
 \label{sec-3}
-\section{Incoherent Dictionary Learning}
+\section{Numerical Experiments}
 \label{sec-4}
-\section{Learning Incoherent Subspaces}
+\section{Discussion}
 \label{sec-5}
-\section{Numerical Experiments}
+\section{Conclusions}
 \label{sec-6}
-\section{Discussion}
-\label{sec-7}
-\section{Conclusions}
-\label{sec-8}
 % Emacs 24.3.1 (Org mode 8.2.4)

### 2013-12-13 manuscript.tex
added definitions and maths packages

@@
-% Created 2013-12-13 Fri 14:51
+% Created 2013-12-13 Fri 16:40
 \documentclass[]{article}
@@
 \tolerance=1000
-\author{Daniele Barchiesi}
+\usepackage{amsfonts,amssymb,amsmath,amsthm,bm}
+\author{Daniele Barchiesi and Mark D. Plumbley}
 \date{\textit{<2013-12-13 Fri>}}
@@
 \maketitle
+\input{/Users/daniele/Dropbox/Career/Research/MyPapers/definitions.tex}
+
 \begin{abstract}
@@
 \label{sec-1}
-\begin{enumerate}
-\item $\square$ prepare a reading list
-\item $\square$ read papers
-\item $\square$ select relevant publications
-\item $\square$ write introduction and summaries
-\end{enumerate}
 \subsection{Problem definition}
 \label{sec-1-1}
-
 \subsection{Previous work and applications}
@@
 Feature transforms have been used to model data in an attempt to enhance the tradeoff between generalisation and discrimination, as described in Section \ref{sec-1}.
-Two of the main feature transform techniques include principal component analysis (PCA) and Fisher's linear discriminant analysis (LDA). This section provides a brief description of their rationale.
 
-Let $\{x_m \in R^N\}$ be a set of vectors containing features extracted from $M$ training signals.
+Two of the most common feature transform techniques include principal component analysis (PCA) \cite{Pearson1901On} and Fisher's linear discriminant analysis (LDA) \cite{Duda1973Pa}.
+This section provides a brief description of their rationale.
 
-\subsubsection{Manifold learning}
-\label{sec-2-1-2}
-
+Let $\{\Vector{x}_m \in \real^N\}$ be a set of vectors containing features extracted from $M$ training signals.
+\subsection{Manifold learning}
+\label{sec-2-2}
 \subsection{Dictionary learning}
-\label{sec-2-2}
+\label{sec-2-3}
 \subsubsection{Dictionary learning for sparse approximation}
-\label{sec-2-2-1}
+\label{sec-2-3-1}
 \subsubsection{Incoherent dictionary learning}
-\label{sec-2-2-2}
+\label{sec-2-3-2}
 \subsubsection{Dictionary learning for classification}
-\label{sec-2-2-3}
+\label{sec-2-3-3}
 
@@
 \label{sec-6}
+
+\bibliographystyle{plain}
+\bibliography{/Users/daniele/Dropbox/Career/Research/MyPapers/bibliography.bib}
 % Emacs 24.3.1 (Org mode 8.2.4)

### 2013-12-14 code/GetFisherIrisDataset.m
added function to load the Fisher's iris dataset from the web

@@ -1,0 +1,8 @@
+function [fea,cat] = GetFisherIris()
+% Gets the fisher iris dataset from the web and returns the following variables
+% - fea: matrix containing sepal length, width, petal length and width along the columns
+% - cat: cell containing the classes
+url = 'http://archive.ics.uci.edu/ml/machine-learning-databases/iris/iris.data';
+s = urlread(url);
+[seplen,sepwid,petlen,petwid,cat] = strread(s,"%f,%f,%f,%f,%s");
+fea = [seplen, sepwid, petlen, petwid];

### 2014-01-06 code/GetFisherIrisDataset.m
added functions for reproducing mlsp figure

@@ -1,8 +1,9 @@
-function [fea,cat] = GetFisherIris()
+function [fea,cat] = GetFisherIrisDataset()
 % Gets the fisher iris dataset from the web and returns the following variables
 % - fea: matrix containing sepal length, width, petal length and width along the columns
 % - cat: cell containing the classes
 url = 'http://archive.ics.uci.edu/ml/machine-learning-databases/iris/iris.data';
 s = urlread(url);
-[seplen,sepwid,petlen,petwid,cat] = strread(s,"%f,%f,%f,%f,%s");
-fea = [seplen, sepwid, petlen, petwid];
+C = textscan(s,'%f,%f,%f,%f,%s');
+fea = [C{1} C{2} C{3} C{4}];
+cat = C{5};

### 2014-01-06 code/ComputeAtomsCategories.m
added code to make paper self-contained (only requires SMALLBox installation)

@@ -1,0 +1,15 @@
+function [cat, C] = ComputeAtomsCategories(X,cat)
+
+if isnumeric(cat), cat = num2str(cat); end
+
+nAto = size(X,1);
+uniCat = unique(cat);
+nCat = length(uniCat);
+%compute a matrix whose (i,j)th element is the contribution of atom i to the sparse approximation of signals in the class j
+C = zeros(nAto,nCat);
+for iCat=1:nCat
+    idx = strcmp(cellstr(cat),char(uniCat(iCat)));
+    C(:,iCat) = sum(abs(X(:,idx)'))/sum(idx);
+end
+
+[~, cat] = max(C,[],2);

### 2014-01-06 code/IPRDis.m

@@ -1,0 +1,261 @@
+function [dic, mus, srs] = IPRDis(dic,mu,Y,X,cat,par)
+%% Unit test
+if nargin==1        %if input argument is a single number, run appropriate unit test
+    feval(strcat('UnitTest',num2str(dic)));
+    return;
+end
+
+%% Parameters and defaults
+if ~exist('par','var') || isempty(par), par = struct; end   %parameters
+
+def.nIte  = 20;     %number of iterations
+def.eps   = 1e-6;   %tolerance level
+def.alpha = .7;     %mix weight
+def.vis   = false;  %visualize results (only for 2 or 3 dimensional datasets)
+def.ver   = false;  %verose output
+
+par = setdefaultoptions(def,par);   %set default parameters
+
+nDim = size(dic,1);  %dimension of atoms of the dictionary
+
+if isnumeric(cat), cat = num2str(cat); end
+if ~isempty(cat)
+    atoCat = ComputeAtomsCategories(X,cat);
+    MU  = @(dic) MUCat(dic,atoCat); %mutual coherence function
+else
+    atoCat = [];
+    MU  = @(dic) max(max(abs((dic'*dic)-diag(diag(dic'*dic))))); %mutual coherence function
+end
+SNR = @(dic) snr(Y,dic*omp(dic'*Y,dic'*dic,par.Tdata));
+
+%% Main algorithm
+dic = normc(dic);               %normalise columns
+
+mus = zeros(par.nIte,1);		%coherence at each iteration
+srs = zeros(par.nIte,1);		%signal to noise ratio at each iteration
+iIter = 1;
+
+if par.vis
+        fprintf(1,'\n--------------------------------\n');
+end
+while iIter<=par.nIte && MU(dic)>mu
+    % calculate snr and coherence
+    mus(iIter) = MU(dic);
+    srs(iIter) = SNR(normc(dic));
+    
+    if par.vis
+        fprintf(1,'Iteration number %u - mu: %f, snr: %f\n', iIter,mus(iIter),srs(iIter));
+    end
+    
+    % calculate gram matrix
+    G = dic'*dic;
+    
+    %project onto the structural constraint set and cumpute convex combination
+    H = ProjectStruct(G,mu,atoCat);
+    alpha = par.alpha;
[…]

### 2014-01-06 code/IPRLearnDisSub.m

@@ -1,0 +1,144 @@
+function [traSubSpa,newTraFea,newTesFea] = IPRLearnDisSub(traFea,tesFea,traCat,par)
+% Learn discriminative subspaces using iterative projections and rotations
+% (IPR) incoherent dictionaty learning algorithm.
+
+if ~nargin, unitTest; return; end
+%% Params and defaults
+if ~exist('par','var') || isempty(par), par=struct; end
+if isnumeric(traCat), traCat = num2str(traCat);     end
+
+uniCat = unique(traCat);                %names of categories
+nCat = length(uniCat);                  %number of categories
+nDim = size(traFea,2);                  %dimension of features
+toolbox   = 'SMALL_incoherentDL';		%dictionary learning toolbox
+
+def.subSpaRan = 50;                     %rank of incoherent subspaces
+def.nAto = nCat*nDim;                   %number of atoms
+def.nIte = 1;                           %number of DL iterations
+
+par = setdefaultoptions(par,def);       %set parameters
+if ~isfield(par,'mu') || isempty(par.mu)
+    par.mu = sqrt((par.nAto-nDim)/(nDim*(par.nAto-1)));
+end
+
+nAto  = par.nAto;                %number of atoms in the dictionary
+perActAto = par.perActAto;   %percentage of active atoms
+subSpaRan = par.subSpaRan;
+
+nActiveAtoms = fix(nDim/100*perActAto); %number of active atoms
+
+
+%% Create SMALL structures for dictionary learning
+SMALL.Problem.b  = traFea';
+SMALL.Problem.b1 = SMALL.Problem.b;
+SMALL.Problem.cat = traCat;
+
+% omp2 sparse representation solver
+ompParam = struct('X',SMALL.Problem.b,'epsilon',1e-6,'maxatoms',nActiveAtoms); %parameters
+solver	 = SMALL_init_solver('ompbox','omp2',ompParam,false); %solver structure
+
+SMALL.DL.toolbox = toolbox;
+SMALL.DL.name = 'ksvd';
+SMALL.DL.profile = true;
+SMALL.DL.param.data = SMALL.Problem.b;
+SMALL.DL.param.Tdata = nActiveAtoms;
+SMALL.DL.param.dictsize = nAto;
+SMALL.DL.param.nIte = par.nIte;
+SMALL.DL.param.memusage = 'high';
[…]

### 2014-01-06 code/LDAFeaturesTransform.m

@@ -1,0 +1,53 @@
+function [newTraFea, newTesFea] = LDAFeaturesTransform(traFea,tesFea,traCat)
+if ~nargin, unitTest; return; end
+
+ldaObj = ClassificationDiscriminant.fit(traFea,traCat);           %fit a LDA object
+Sb = ldaObj.BetweenSigma;                   %Between classes scatter matrix
+Sw = ldaObj.Sigma;                          %Within classes scatter matrix
+St = Sb + Sw;                               %Total scatter matrix
+
+J = @(W) trace(pinv(W'*St*W)*(W'*Sb*W));    %as a function of projection matrix
+
+[D,X] = eig(pinv(St)*Sb);                   %compute evd
+idx = diag(real(X))>1e-8;                         %find non-zero eigenvalues
+W = real(D(:, idx));                              %select corresponding eigenvectors
+
+newTraFea = (W*pinv(W)*traFea')';           %project data onto subspace spanned by eigenvectors
+newTesFea = (W*pinv(W)*tesFea')';
+
+
+function unitTest
+clear, clc, close all
+load fisheriris
+nDim = 4;
+nObs = 150;
+traFea = meas(1:nObs,1:nDim);
+tesFea = traFea;
+traCat = species(1:nObs);
+traFea = traFea-repmat(mean(traFea),nObs,1);
+gscatter(traFea(:,1), traFea(:,2), traCat,'rgb','osd');
+xlabel('Sepal length');
+ylabel('Sepal width');
+
+[newTraFea,newTesFea] = LDAFeaturesTransform(traFea,tesFea,traCat);
+
+figure
+gscatter(newTraFea(:,1), newTraFea(:,2), traCat,'rgb','osd');
+xlabel('Sepal length');
+ylabel('Sepal width');
+
+J1 = @(v) trace(pinv(ClassificationDiscriminant.fit(v,traCat).Sigma)*ClassificationDiscriminant.fit(v,traCat).BetweenSigma);
+J2 = @(v) trace(ClassificationDiscriminant.fit(v,traCat).BetweenSigma)/trace(ClassificationDiscriminant.fit(v,traCat).Sigma);
+J3 = @(v) log(abs(det(ClassificationDiscriminant.fit(v,traCat).BetweenSigma))) - ClassificationDiscriminant.fit(v,traCat).LogDetSigma;
+
[…]

### 2014-01-06 code/NormalizeFeatures.m

@@ -1,0 +1,11 @@
+function [traFea, tesFea] = NormalizeFeatures(traFea,tesFea)
+m = mean(traFea,1);
+s = std(traFea,[],1);
+for iFea=1:size(traFea,2)
+    traFea(:,iFea) = (traFea(:,iFea)-m(iFea))./max(s(iFea),realmin);
+end
+if exist('tesFea','var')
+    for iFea=1:size(tesFea,2)
+        tesFea(:,iFea) = (tesFea(:,iFea)-m(iFea))./s(iFea);
+    end
+end

### 2014-01-06 code/PCAFeaturesTransform.m

@@ -1,0 +1,38 @@
+function [newTraFea, newTesFea] = PCAFeaturesTransform(traFea,tesFea,~,thresh)
+if ~nargin, unitTest; return; end
+if ~exist('thresh','var') || isempty(thresh), thresh = 95; end
+%compute principal components and weights
+[D,~,~,~,explained] = pca(traFea);
+X = traFea*D';
+%find a number of pc that explain 95% of the variance in the data
+thrIdx = find(cumsum(explained)>=thresh,1);
+%transformed training features
+newTraFea = X(:,1:thrIdx)*D(1:thrIdx,1:thrIdx);
+%transform test features using the dictionary learned on the
+%training features
+XTes = tesFea*D';                   %pca coefficients
+newTesFea = XTes(:,1:thrIdx)*D(1:thrIdx,1:thrIdx);
+
+function unitTest
+clear, clc, close all
+load fisheriris
+nDim = 4;
+nObs = 150;
+traFea = meas(1:nObs,1:nDim);
+tesFea = traFea;
+traCat = species(1:nObs);
+traFea = traFea-repmat(mean(traFea),nObs,1);
+gscatter(traFea(:,1), traFea(:,2), traCat,'rgb','osd');
+xlabel('Sepal length');
+ylabel('Sepal width');
+
+[newTraFea,newTesFea] = PCAFeaturesTransform(traFea,tesFea);
+size(newTraFea,2);
+figure
+if size(newTraFea,2)<2
+    gscatter(newTraFea(:,1), newTraFea(:,1), traCat,'rgb','osd');
+else
+    gscatter(newTraFea(:,1), newTraFea(:,2), traCat,'rgb','osd');
+end
+xlabel('Sepal length');
+ylabel('Sepal width');

### 2014-01-06 code/SMALL_incoherentDL.m

@@ -1,0 +1,181 @@
+function DL=SMALL_incoherentDL(Problem, DL)
+
+% determine which solver is used for sparse representation %
+
+solver = DL.param.solver;
+
+% determine which type of udate to use ('KSVD', 'MOD', 'ols' or 'mailhe') %
+
+typeUpdate = DL.name;
+
+sig = Problem.b;
+
+% determine dictionary size %
+
+if (isfield(DL.param,'initdict'))
+    if (any(size(DL.param.initdict)==1) && all(iswhole(DL.param.initdict(:))))
+        dictsize = length(DL.param.initdict);
+    else
+        dictsize = size(DL.param.initdict,2);
+    end
+end
+if (isfield(DL.param,'dictsize'))    % this superceedes the size determined by initdict
+    dictsize = DL.param.dictsize;
+end
+
+if (size(sig,2) < dictsize)
+    error('Number of training signals is smaller than number of atoms to train');
+end
+
+
+% initialize the dictionary %
+
+% todo: check second if statement
+if (isfield(DL.param,'initdict')) && ~isempty(DL.param.initdict);
+    if (any(size(DL.param.initdict)==1) && all(iswhole(DL.param.initdict(:))))
+        dico = sig(:,DL.param.initdict(1:dictsize));
+    else
+        if (size(DL.param.initdict,1)~=size(sig,1) || size(DL.param.initdict,2)<dictsize)
+            error('Invalid initial dictionary');
+        end
+        dico = DL.param.initdict(:,1:dictsize);
+    end
+else
+    data_ids = find(colnorms_squared(sig) > 1e-6);   % ensure no zero data elements are chosen
+    perm = randperm(length(data_ids));
+    dico = sig(:,data_ids(perm(1:dictsize)));
+end
+
+% flow: 'sequential' or 'parallel'. If sequential, the residual is updated
+% after each atom update. If parallel, the residual is only updated once
+% the whole dictionary has been computed. Sequential works better, there
+% may be no need to implement parallel. Not used with MOD.
+
+if isfield(DL.param,'flow')
+    flow =  DL.param.flow;
+else
+    flow = 'sequential';
+end
+
+% learningRate. If the type is 'ols', it is the descent step of
+% the gradient (typical choice: 0.1). If the type is 'mailhe', the
+% descent step is the optimal step*rho (typical choice: 1, although 2
[…]

### 2014-01-06 code/TestTransforms.m

@@ -1,0 +1,132 @@
+function res = TestTransforms(traFea,tesFea,traCat,tesCat,par)
+
+addpath(genpath(pwd));
+reset(RandStream.getGlobalStream)
+
+%% Parameters and defaults
+if ~exist('par','var') || isempty(par), par = struct; end
+
+def.nAto = 10;
+def.subSpaRan = 1;
+def.perActAto = 50;
+def.nKNNs = 5;
+
+par = setdefaultoptions(par,def);
+
+if ~exist('traFea','var') || isempty(traFea) || ~exist('tesFea','var') || isempty(tesFea)...
+        || ~exist('traCat','var') || isempty(traCat) || ~exist('tesCat','var') || isempty(tesCat)
+    [meas,species] = GetFisherIrisDataset;
+    nDim = 3;
+    nObs = 150;
+    v = randperm(nObs);
+    traFea = meas(v(1:100),1:nDim);
+    tesFea = meas(v(101:150),1:nDim);
+    traCat = species(v(1:100));
+    tesCat = species(v(101:150));
+    feaMea = mean(traFea);
+    feaStd = std(traFea);
+    traFea = (traFea-repmat(feaMea,100,1))./repmat(feaStd,100,1);
+    tesFea = (tesFea-repmat(feaMea,50,1))./repmat(feaStd,50,1);
+end
+
+%% Classify using linear discriminant analysis
+cvPar = cvpartition([traCat;tesCat],'kfold',5);                 %create 5fold partition
+
+conMatFun = @(xTra,yTra,xTes,yTes) (confusionmat(yTes,...               %Compute confusion matrix
+    nominal(knnclassify(xTes,xTra,yTra,par.nKNNs))));
+conMat = crossval(conMatFun,[traFea;tesFea],nominal([traCat;tesCat]),'partition',cvPar);
+
+nCat = length(unique(nominal([traCat;tesCat])));
+conMat = reshape(sum(conMat),nCat,nCat);
+mcr = sum(sum(conMat-diag(diag(conMat))))/sum(sum(conMat));
+
+%% Test PCA feature transform
+[pcaTraFea, pcaTesFea] = PCAFeaturesTransform(traFea,tesFea);
+
+pcaConMat = crossval(conMatFun,[pcaTraFea;pcaTesFea],nominal([traCat;tesCat]),'partition',cvPar);
[…]

### 2014-01-06 code/setdefaultoptions.m

@@ -1,0 +1,27 @@
+function options = setdefaultoptions(options, defaults)
+% Assign the defaults value to any field not specified in the strucy
+% options
+%
+%  options = ManageOptions(defaults, options)
+%
+% INPUT
+%  options (struct) the current options
+%  defaults (struct) the default options
+%
+% OUTPUT
+%  options (struct) the updated options
+%
+
+if nargin<2 || isempty(options)
+    options = defaults;
+else
+    fields = fieldnames(options);
+    defaultFields = fieldnames(defaults);
+    numFields = length(defaultFields);
+    for i=1:numFields
+        if ~isfield(options, defaultFields(i))
+            value = getfield(defaults, char(defaultFields(i)));
+            options = setfield(options, char(defaultFields(i)), value);
+        end
+    end
+end

### 2014-01-06 code/PCAFeaturesTransform.m
implemented supervised pca algorithm

@@ -14,7 +14,7 @@ XTes = tesFea*D';                   %pca coefficients
 newTesFea = XTes(:,1:thrIdx)*D(1:thrIdx,1:thrIdx);
 
 function unitTest
-clear, clc, close all
+clear, clc
 load fisheriris
 nDim = 4;
 nObs = 150;
@@ -22,6 +22,7 @@ traFea = meas(1:nObs,1:nDim);
 tesFea = traFea;
 traCat = species(1:nObs);
 traFea = traFea-repmat(mean(traFea),nObs,1);
+figure
 gscatter(traFea(:,1), traFea(:,2), traCat,'rgb','osd');
 xlabel('Sepal length');
 ylabel('Sepal width');
@@ -34,5 +35,5 @@ if size(newTraFea,2)<2
 else
     gscatter(newTraFea(:,1), newTraFea(:,2), traCat,'rgb','osd');
 end
-xlabel('Sepal length');
-ylabel('Sepal width');
+xlabel('1st principal component');
+ylabel('2nd principal component');

### 2014-01-06 code/SPCAFeaturesTransform.m

@@ -1,0 +1,62 @@
+function [newTraFea, newTesFea] = SPCAFeaturesTransform(traFea,tesFea,cat,thresh)
+%supervised pca implementation. For details, please refer to Barshan et al.
+%, Pattern Recognition, 44(2011) 1357--1371
+
+if ~nargin, unitTest; return; end
+if ~exist('thresh','var') || isempty(thresh), thresh = 100; end
+
+%compute centering matrix
+X = traFea';
+n = size(X,2); %number of observations
+H = eye(n) - ones(n,1)*ones(n,1)'/n;
+
+%compute objective matrix Q
+if ~isnumeric(cat), cat = NumericEquivalent(cat); end %if category is not numeric, compute the equivalent
+L = cat*cat'; %kernel of responses (classes)
+Q = X*H*L*H*X';
+
+%compute principal components and weights
+[D,~,explained] = pcacov(Q); %use pcacov because Q is a covariance matrix
+
+%find a number of pc that explain thresh% of the variance in the data
+thrIdx = find(cumsum(explained)>=thresh,1);
+
+%transformed training features
+U = D(:,1:thrIdx);
+newTraFea = X'*U; %return n observations along the rows of matrix
+newTesFea = tesFea*U;
+
+function newCat = NumericEquivalent(cat)
+%takes a cell of string categories and returns a vector of equivalent
+%integer values. Example: {'a','b','c','b','a'} => [1,2,3,2,1].
+nObs = length(cat);
+uniCat = unique(cat);
+nCat = length(uniCat);
+newCat = zeros(nObs,1);
+for i=1:nCat
+    newCat(strcmp(cat,uniCat(i))) = i-1;
+end
+
+function unitTest
+clear, clc
+load fisheriris
+nDim = 4;
+nObs = 150;
+traFea = meas(1:nObs,1:nDim);
+tesFea = traFea;
+traCat = species(1:nObs);
+traFea = traFea-repmat(mean(traFea),nObs,1);
+figure
+gscatter(traFea(:,1), traFea(:,2), traCat,'rgb','osd');
+xlabel('Sepal length');
+ylabel('Sepal width');
+
+newTraFea = SPCAFeaturesTransform(traFea,tesFea,traCat);
+figure
+if size(newTraFea,2)<2
+    gscatter(newTraFea(:,1), newTraFea(:,1), traCat,'rgb','osd');
+else
[…]

### 2014-01-07 code/SPCAFeaturesTransform.m
modified TestTransforms to include spca and distinction between training and test points

@@ -18,45 +18,68 @@ Q = X*H*L*H*X';
 %compute principal components and weights
 [D,~,explained] = pcacov(Q); %use pcacov because Q is a covariance matrix
 
-%find a number of pc that explain thresh% of the variance in the data
-thrIdx = find(cumsum(explained)>=thresh,1);
+%find a number of pc that explain thresh% of the variance in the data or
+%min 2d
+thrIdx = 1:find(thresh-cumsum(explained)>=100*eps); %use this to avoid machine errors
+if length(thrIdx) < 2, thrIdx = [1,2]; end
 
 %transformed training features
-U = D(:,1:thrIdx);
+U = D(:,thrIdx);
 newTraFea = X'*U; %return n observations along the rows of matrix
 newTesFea = tesFea*U;
 
 function newCat = NumericEquivalent(cat)
 %takes a cell of string categories and returns a vector of equivalent
 %integer values. Example: {'a','b','c','b','a'} => [1,2,3,2,1].
+cat = cellstr(cat); %transform to string in case of nominal categories
 nObs = length(cat);
 uniCat = unique(cat);
 nCat = length(uniCat);
 newCat = zeros(nObs,1);
 for i=1:nCat
-    newCat(strcmp(cat,uniCat(i))) = i-1;
+    newCat(strcmp(cellstr(cat),uniCat(i))) = i-1;
 end
 
 function unitTest
 clear, clc
+traObsPer = 0.7;
 load fisheriris
 nDim = 4;
 nObs = 150;
-traFea = meas(1:nObs,1:nDim);
-tesFea = traFea;
-traCat = species(1:nObs);
-traFea = traFea-repmat(mean(traFea),nObs,1);
-figure
-gscatter(traFea(:,1), traFea(:,2), traCat,'rgb','osd');
+iObs = randperm(nObs);
+traFea = meas(iObs(1:floor(traObsPer*nObs)),1:nDim);
+tesFea = meas(iObs(floor(traObsPer*nObs)+1:end),1:nDim);
+traCat = species(iObs(1:floor(traObsPer*nObs)));
+tesCat = species(iObs(floor(traObsPer*nObs)+1:end));
+
+% Center features
+mFea = mean(traFea);
+traFea = traFea-repmat(mFea,floor(traObsPer*nObs),1);
[…]

### 2014-01-07 code/TestTransforms.m

@@ -1,79 +1,81 @@
 function res = TestTransforms(traFea,tesFea,traCat,tesCat,par)
 
-addpath(genpath(pwd));
-reset(RandStream.getGlobalStream)
+addpath(genpath(pwd)); %add path to matlab search path
+reset(RandStream.getGlobalStream) %reset random stream to reproduce results exactly
 
 %% Parameters and defaults
 if ~exist('par','var') || isempty(par), par = struct; end
 
-def.nAto = 10;
-def.subSpaRan = 1;
-def.perActAto = 50;
-def.nKNNs = 5;
+def.nAto = 10; %number of atoms
+def.subSpaRan = 1; %subspace range
+def.perActAto = 50; %percentage of active atoms
+def.nKNNs = 5; %number of nearest neighbours for classification
 
 par = setdefaultoptions(par,def);
 
 if ~exist('traFea','var') || isempty(traFea) || ~exist('tesFea','var') || isempty(tesFea)...
         || ~exist('traCat','var') || isempty(traCat) || ~exist('tesCat','var') || isempty(tesCat)
-    [meas,species] = GetFisherIrisDataset;
-    nDim = 3;
-    nObs = 150;
-    v = randperm(nObs);
-    traFea = meas(v(1:100),1:nDim);
-    tesFea = meas(v(101:150),1:nDim);
-    traCat = species(v(1:100));
-    tesCat = species(v(101:150));
-    feaMea = mean(traFea);
-    feaStd = std(traFea);
-    traFea = (traFea-repmat(feaMea,100,1))./repmat(feaStd,100,1);
-    tesFea = (tesFea-repmat(feaMea,50,1))./repmat(feaStd,50,1);
+    [meas,species] = GetFisherIrisDataset; %get data from iris dataset
+    nDim = 3; %data has 4 dimensions
+    nObs = 150;  %and 150 observations
+    v = randperm(nObs); %select random permutation of observations
+    perTraObs = 70; %percentage of observations used for training
+    nTraObs = floor(nObs*perTraObs/100); %number of training observations
+    traFea = meas(v(1:nTraObs),1:nDim); %training data
[…]

### 2014-01-08 code/IPRLearnDisSub.m
fixed visualization problem for IPR

@@ -16,10 +16,8 @@ def.subSpaRan = 50;                     %rank of incoherent subspaces
 def.nAto = nCat*nDim;                   %number of atoms
 def.nIte = 1;                           %number of DL iterations
 
+def.mu = sqrt((par.nAto-nDim)/(nDim*(par.nAto-1))); %mutual coherence
 par = setdefaultoptions(par,def);       %set parameters
-if ~isfield(par,'mu') || isempty(par.mu)
-    par.mu = sqrt((par.nAto-nDim)/(nDim*(par.nAto-1)));
-end
 
 nAto  = par.nAto;                %number of atoms in the dictionary
 perActAto = par.perActAto;   %percentage of active atoms
@@ -62,11 +60,11 @@ Xtra = solver.solution;
 
 [q,C] = ComputeAtomsCategories(Xtra,traCat);  %atom's contributions to categories
 for iCat=1:nCat     %for every category
-    [~,ind] = sort(C(:,iCat),'descend');    %sort atoms according to contributions to category
-    traSubSpa{iCat} = D(:,ind(1:subSpaRan));%select subset of atoms in the dictionary as subspace
+    [~,idx] = sort(C(:,iCat),'descend');    %sort atoms according to contributions to category
+    traSubSpa{iCat} = D(:,idx(1:subSpaRan));%select subset of atoms in the dictionary as subspace
     c=1;                                    %counter
     while rank(traSubSpa{iCat})<subSpaRan   %if and until rank of subspace is less than predefined rank
-        traSubSpa{iCat} = [traSubSpa{iCat}, D(:,ind(subSpaRan+c))];  %add next atom
+        traSubSpa{iCat} = [traSubSpa{iCat}, D(:,idx(subSpaRan+c))];  %add next atom
         c = c+1;
     end
 end
@@ -80,10 +78,10 @@ for iCat=1:nCat
     for jCat=1:nCat
         catIdx{iCat} = strcmp(cellstr(traCat),char(uniCat(iCat)));     %indexes corresponding to iCat category
         tem = Project(traSubSpa{jCat},traFea(catIdx{iCat},:)')';       %Projection of signals of iCat catogory onto jCat subspace
-        dis(iCat,jCat) = norm(traFea(catIdx{iCat},:)-tem,'fro');       %distance
+        dis(iCat,jCat) = norm(traFea(catIdx{iCat},:)-tem,'fro');       %distance between data belonging to class iCat and their projection onto subspace jCat
     end
[…]

### 2014-01-08 code/PCAFeaturesTransform.m

@@ -1,17 +1,23 @@
-function [newTraFea, newTesFea] = PCAFeaturesTransform(traFea,tesFea,~,thresh)
+function [newTraFea, newTesFea] = PCAFeaturesTransform(traFea,tesFea,~,par)
 if ~nargin, unitTest; return; end
-if ~exist('thresh','var') || isempty(thresh), thresh = 95; end
+if ~exist('par','var') || isempty(par), par=struct; end
+def.thresh = 95;
+def.subSpaRan = [];
+par = setdefaultoptions(par,def);
+
 %compute principal components and weights
 [D,~,~,~,explained] = pca(traFea);
-X = traFea*D';
+if par.subSpaRan
+    D = D(:,1:par.subSpaRan);
+else
 %find a number of pc that explain 95% of the variance in the data
-thrIdx = find(cumsum(explained)>=thresh,1);
+    thrIdx = find(cumsum(explained)>=par.thresh,1);
+    D = D(:,1:thrIdx);
+end
+Project = @(Phi,x) Phi*pinv(Phi)*x;
 %transformed training features
-newTraFea = X(:,1:thrIdx)*D(1:thrIdx,1:thrIdx);
-%transform test features using the dictionary learned on the
-%training features
-XTes = tesFea*D';                   %pca coefficients
-newTesFea = XTes(:,1:thrIdx)*D(1:thrIdx,1:thrIdx);
+newTraFea = Project(D,traFea);
+newTesFea = Project(D,tesFea);
 
 function unitTest
 clear, clc

### 2014-01-08 code/SPCAFeaturesTransform.m

@@ -1,9 +1,12 @@
-function [newTraFea, newTesFea] = SPCAFeaturesTransform(traFea,tesFea,cat,thresh)
+function [newTraFea, newTesFea] = SPCAFeaturesTransform(traFea,tesFea,cat,par)
 %supervised pca implementation. For details, please refer to Barshan et al.
 %, Pattern Recognition, 44(2011) 1357--1371
 
 if ~nargin, unitTest; return; end
-if ~exist('thresh','var') || isempty(thresh), thresh = 100; end
+if ~exist('par','var') || isempty(par), par = struct; end
+def.thresh = 100;
+def.subSpaRan = [];
+par = setdefaultoptions(par,def);
 
 %compute centering matrix
 X = traFea';
@@ -18,13 +21,17 @@ Q = X*H*L*H*X';
 %compute principal components and weights
 [D,~,explained] = pcacov(Q); %use pcacov because Q is a covariance matrix
 
-%find a number of pc that explain thresh% of the variance in the data or
-%min 2d
-thrIdx = 1:find(thresh-cumsum(explained)>=100*eps); %use this to avoid machine errors
-if length(thrIdx) < 2, thrIdx = [1,2]; end
-
-%transformed training features
-U = D(:,thrIdx);
+if par.subSpaRan
+    U = D(:,1:par.subSpaRan);
+else
+    %find a number of pc that explain thresh% of the variance in the data or
+    %min 2d
+    thrIdx = 1:find(thresh-cumsum(explained)>=100*eps); %use this to avoid machine errors
+    if length(thrIdx) < 2, thrIdx = [1,2]; end
+    
+    %transformed training features
+    U = D(:,thrIdx);
+end
 newTraFea = X'*U; %return n observations along the rows of matrix
 newTesFea = tesFea*U;
 

### 2014-01-08 code/TestTransforms.m

@@ -10,13 +10,14 @@ def.nAto = 10; %number of atoms
 def.subSpaRan = 1; %subspace range
 def.perActAto = 50; %percentage of active atoms
 def.nKNNs = 5; %number of nearest neighbours for classification
+def.visu = false;
 
 par = setdefaultoptions(par,def);
 
 if ~exist('traFea','var') || isempty(traFea) || ~exist('tesFea','var') || isempty(tesFea)...
         || ~exist('traCat','var') || isempty(traCat) || ~exist('tesCat','var') || isempty(tesCat)
     [meas,species] = GetFisherIrisDataset; %get data from iris dataset
-    nDim = 3; %data has 4 dimensions
+    nDim = 4; %data has 4 dimensions
     nObs = 150;  %and 150 observations
     v = randperm(nObs); %select random permutation of observations
     perTraObs = 70; %percentage of observations used for training
@@ -63,7 +64,7 @@ ldaConMat = reshape(sum(ldaConMat),nCat,nCat);
 ldaMcr = sum(sum(ldaConMat-diag(diag(ldaConMat))))/sum(sum(ldaConMat));
 
 %% Test IPR feature transform
-[~, iprTraFea, iprTesFea] = IPRLearnDisSub(traFea,tesFea,traCat,par);
+[subs, iprTraFea, iprTesFea] = IPRLearnDisSub(traFea,tesFea,traCat,par);
 conMatFun = @(traFea,traCat,tesFea,tesCat)IPRTransformAndClassify(traFea,traCat,tesFea,tesCat,par);
 iprConMat = crossval(conMatFun,[iprTraFea;iprTesFea],nominal([traCat;tesCat]),'partition',cvPar);
 iprConMat = reshape(sum(iprConMat),nCat,nCat);
@@ -74,44 +75,45 @@ iprMcr = sum(sum(iprConMat-diag(diag(iprConMat))))/sum(sum(iprConMat));
 % J1 = @(fea,cat) trace(pinv(ClassificationDiscriminant.fit(fea,cat).Sigma)*ClassificationDiscriminant.fit(fea,cat).BetweenSigma);
 % J2 = @(fea,cat) trace(ClassificationDiscriminant.fit(fea,cat).BetweenSigma)/trace(ClassificationDiscriminant.fit(fea,cat).Sigma);
 % J3 = @(fea,cat) log(abs(det(ClassificationDiscriminant.fit(fea,cat).BetweenSigma))) - ClassificationDiscriminant.fit(fea,cat).LogDetSigma;
-
[…]

### 2014-01-08 code/GetBalanceDataset.m
added functions to get datasets and function to test classification with different dimensions

@@ -1,0 +1,7 @@
+function [fea,cat] = GetBalanceDataset()
+% Gets the fisher balance dataset from the web (see http://archive.ics.uci.edu/ml/datasets/Balance+Scale)
+url = 'http://archive.ics.uci.edu/ml/machine-learning-databases/balance-scale/balance-scale.data';
+s = urlread(url);
+C = textscan(s,'%s %f %f %f %f','Delimiter',',');
+fea = [C{2:5}];
+cat = C{1};

### 2014-01-08 code/GetColonCancerDataset.m

@@ -1,0 +1,22 @@
+function [fea,cat] = GetColonCancerDataset()
+% Gets the colon cancer dataset from the web
+url = 'http://genomics-pubs.princeton.edu/oncology/affydata/I2000.html';
+s = urlread(url);
+s = regexprep(s,'<.*?>','');
+format = repmat('%f',1,62);
+C = textscan(s(26:end),format,'CollectOutput',true);
+fea = C{1}';
+
+catUrl = 'http://genomics-pubs.princeton.edu/oncology/affydata/tissues.html';
+s = urlread(catUrl);
+s = regexprep(s,'<.*?>','');
+format = '%f';
+C = textscan(s(26:end),format);
+cat = cell(size(C{1},1),1);
+for i=1:length(cat)
+    if C{1}(i)<0
+        cat{i} = 'tumor';
+    else
+        cat{i} = 'normal';
+    end
+end

### 2014-01-08 code/GetParkinsonsDataset.m

@@ -1,0 +1,8 @@
+function [fea,cat] = GetParkinsonsDataset()
+% Gets the Parkinsons dataset from the web
+url = 'http://archive.ics.uci.edu/ml/machine-learning-databases/parkinsons/parkinsons.data';
+s = urlread(url);
+format = '%s %f %f %f %f %f %f %f %f %f %f %f %f %f %f %f %f %s %f %f %f %f %f %f';
+C = textscan(s(230:end),format,'Delimiter',',');
+fea = [C{2:17} C{19:24}];
+cat = C{18};

### 2014-01-08 code/GetSonarDataset.m

@@ -1,0 +1,10 @@
+function [fea,cat] = GetSonarDataset()
+% Gets the sonar dataset from the web
+url = 'http://archive.ics.uci.edu/ml/machine-learning-databases/undocumented/connectionist-bench/sonar/sonar.all-data';
+s = urlread(url);
+format = [];
+for i=1:60, format = [format, sprintf('%%f ')]; end
+format = [format, '%s'];
+C = textscan(s,format,'Delimiter',',');
+fea = [C{1:60}];
+cat = C{61};

### 2014-01-08 code/IPRLearnDisSub.m

@@ -37,7 +37,7 @@ solver	 = SMALL_init_solver('ompbox','omp2',ompParam,false); %solver structure
 
 SMALL.DL.toolbox = toolbox;
 SMALL.DL.name = 'ksvd';
-SMALL.DL.profile = true;
+SMALL.DL.profile = false;
 SMALL.DL.param.data = SMALL.Problem.b;
 SMALL.DL.param.Tdata = nActiveAtoms;
 SMALL.DL.param.dictsize = nAto;
@@ -89,7 +89,7 @@ end
 newTesFea = nan(size(tesFea));
 temp = nan(size(tesFea,2),nCat);
 dis = nan(nCat,1);
-for iTesFea=1:length(tesFea)        %for every data point in the test set
+for iTesFea=1:size(tesFea,1)        %for every data point in the test set
     for iCat=1:nCat                 %for every caegory
         temp(:,iCat) = Project(traSubSpa{iCat}, tesFea(iTesFea,:)');   %projection of datapoint onto subspace
         dis(iCat) = norm(tesFea(iTesFea,:)'-temp(:,iCat));    %distance between original datapoint and projection

### 2014-01-08 code/PCAFeaturesTransform.m

@@ -1,4 +1,4 @@
-function [newTraFea, newTesFea] = PCAFeaturesTransform(traFea,tesFea,~,par)
+function [newTraFea, newTesFea] = PCAFeaturesTransform(traFea,tesFea,par)
 if ~nargin, unitTest; return; end
 if ~exist('par','var') || isempty(par), par=struct; end
 def.thresh = 95;
@@ -16,8 +16,8 @@ else
 end
 Project = @(Phi,x) Phi*pinv(Phi)*x;
 %transformed training features
-newTraFea = Project(D,traFea);
-newTesFea = Project(D,tesFea);
+newTraFea = Project(D,traFea')';
+newTesFea = Project(D,tesFea')';
 
 function unitTest
 clear, clc

### 2014-01-08 code/SMALL_incoherentDL.m

@@ -23,11 +23,6 @@ if (isfield(DL.param,'dictsize'))    % this superceedes the size determined by i
     dictsize = DL.param.dictsize;
 end
 
-if (size(sig,2) < dictsize)
-    error('Number of training signals is smaller than number of atoms to train');
-end
-
-
 % initialize the dictionary %
 
 % todo: check second if statement
@@ -43,7 +38,16 @@ if (isfield(DL.param,'initdict')) && ~isempty(DL.param.initdict);
 else
     data_ids = find(colnorms_squared(sig) > 1e-6);   % ensure no zero data elements are chosen
     perm = randperm(length(data_ids));
-    dico = sig(:,data_ids(perm(1:dictsize)));
+    if (size(sig,2) < dictsize) %if the number of atoms is larger than the number of signals
+        dico = zeros(size(sig,1),dictsize);
+        for iAto=1:dictsize %for every atom
+            ato = sig(:,data_ids(randperm(length(data_ids),1))); %select a random signal
+            ato = ato + randn(size(ato))*norm(ato)*0.1; %perturb the atom with Gaussian noise
+            dico(:,iAto) = ato;
+        end
+    else
+        dico = sig(:,data_ids(perm(1:dictsize)));
+    end
 end
 
 % flow: 'sequential' or 'parallel'. If sequential, the residual is updated

### 2014-01-08 code/SPCAFeaturesTransform.m

@@ -26,7 +26,7 @@ if par.subSpaRan
 else
     %find a number of pc that explain thresh% of the variance in the data or
     %min 2d
-    thrIdx = 1:find(thresh-cumsum(explained)>=100*eps); %use this to avoid machine errors
+    thrIdx = 1:find(par.thresh-cumsum(explained)>=100*eps); %use this to avoid machine errors
     if length(thrIdx) < 2, thrIdx = [1,2]; end
     
     %transformed training features

### 2014-01-08 code/TestClassSubSpaRan.m

@@ -1,0 +1,27 @@
+function TestClassSubSpaRan
+subSpaRanks = 1:4;
+datasets    = {'coloncancer'};
+datfun = @GetColonCancerDataset;
+par.visu = false;
+for iDat=1:length(datasets)
+    fprintf('\nobtaining dataset %s... ',datasets{iDat});
+    [fea,cat] = datfun();
+    if ~isfloat(fea), fea = double(fea); end     %transform features to float if needed
+    fprintf('done\n');
+    for iSSR = 1:length(subSpaRanks)
+        fprintf('testing transforms using subspaces of rank %d... ',subSpaRanks(iSSR));
+        par.subSpaRan = iSSR;
+        res(iSSR) = TestTransforms(fea,cat,par);
+        fprintf('done\n');
+    end
+    mcrs = [res(:).mcr]';
+    figure, PlotMCRs(mcrs(:,2:end));
+end
+
+function PlotMCRs(mcrs)
+methodNames = {'PCA','SPCA','LDA','IPR'};
+plot(mcrs,'o-','MarkerSize',10)
+legend(methodNames);
+xlabel('Sub-space rank');
+ylabel('misclassification ratio');
+grid on

### 2014-01-08 code/TestTransforms.m

@@ -1,4 +1,4 @@
-function res = TestTransforms(traFea,tesFea,traCat,tesCat,par)
+function res = TestTransforms(fea,cat,par)
 
 addpath(genpath(pwd)); %add path to matlab search path
 reset(RandStream.getGlobalStream) %reset random stream to reproduce results exactly
@@ -6,7 +6,7 @@ reset(RandStream.getGlobalStream) %reset random stream to reproduce results exac
 %% Parameters and defaults
 if ~exist('par','var') || isempty(par), par = struct; end
 
-def.nAto = 10; %number of atoms
+def.nAto = 2*size(fea,2); %number of atoms (twice overcomplete dictionary)
 def.subSpaRan = 1; %subspace range
 def.perActAto = 50; %percentage of active atoms
 def.nKNNs = 5; %number of nearest neighbours for classification
@@ -14,23 +14,28 @@ def.visu = false;
 
 par = setdefaultoptions(par,def);
 
-if ~exist('traFea','var') || isempty(traFea) || ~exist('tesFea','var') || isempty(tesFea)...
-        || ~exist('traCat','var') || isempty(traCat) || ~exist('tesCat','var') || isempty(tesCat)
-    [meas,species] = GetFisherIrisDataset; %get data from iris dataset
+if ~exist('fea','var') || isempty(fea) || ~exist('cat','var') || isempty(cat)
+    [fea,cat] = GetFisherIrisDataset; %get data from iris dataset
     nDim = 4; %data has 4 dimensions
-    nObs = 150;  %and 150 observations
-    v = randperm(nObs); %select random permutation of observations
-    perTraObs = 70; %percentage of observations used for training
-    nTraObs = floor(nObs*perTraObs/100); %number of training observations
-    traFea = meas(v(1:nTraObs),1:nDim); %training data
-    tesFea = meas(v(nTraObs+1:end),1:nDim); %test data
-    traCat = species(v(1:nTraObs)); %training categories
-    tesCat = species(v(nTraObs+1:end)); %test categories
-    feaMea = mean(traFea); %mean of training categories
[…]

### 2014-01-08 code/SMALL/SMALL_init_solver.m
added functions to make scripts completely idependent (don't require smallbox)

@@ -1,0 +1,45 @@
+function solver = SMALL_init_solver(toolbox, name, param, profile)
+%%   Function initialise SMALL structure for sparse representation.
+%   Optional input variables:
+%       toolbox - name of Dictionary Learning toolbox you want to use
+%       name    - name of the algorithm from above toolbox
+%       param   - parameters you want to set
+
+%
+%   Centre for Digital Music, Queen Mary, University of London.
+%   This file copyright 2010 Ivan Damnjanovic.
+%
+%   This program is free software; you can redistribute it and/or
+%   modify it under the terms of the GNU General Public License as
+%   published by the Free Software Foundation; either version 2 of the
+%   License, or (at your option) any later version.  See the file
+%   COPYING included with this distribution for more information.
+%
+%%
+
+if ~ exist( 'toolbox', 'var' ) || isempty(toolbox) 
+    solver.toolbox = []; 
+else
+    solver.toolbox = toolbox;
+end
+if ~ exist( 'name', 'var' ) || isempty(name) 
+    solver.name = [];
+else
+    solver.name = name;
+end
+if ~ exist( 'param', 'var' ) || isempty(param) 
+    solver.param = [];
+else
+    solver.param = param;
+end
+if ~ exist( 'profile', 'var' ) || isempty(profile) 
+    solver.profile = 1;
+else
+    solver.profile = profile;
+end
+solver.add_constraints = 0;
+solver.solution = [];
+solver.reconstructed = [];
+solver.time = [];
+
+end

### 2014-01-08 code/SMALL/SMALL_learn.m

@@ -1,0 +1,50 @@
+function DL = SMALL_learn(Problem,DL)
+%% SMALL Dictionary Learning
+%
+%   Function gets as input Problem and Dictionary Learning (DL) structures
+%   In Problem structure field b with the training set needs to be defined
+%   In DL fields with name of the toolbox and solver, and parameters file
+%   for particular dictionary learning technique needs to be present.
+%
+%   Outputs are Learned dictionary and time spent as a part of DL structure
+
+%
+%   Centre for Digital Music, Queen Mary, University of London.
+%   This file copyright 2009 Ivan Damnjanovic.
+%
+%   This program is free software; you can redistribute it and/or
+%   modify it under the terms of the GNU General Public License as
+%   published by the Free Software Foundation; either version 2 of the
+%   License, or (at your option) any later version.  See the file
+%   COPYING included with this distribution for more information.
+%%
+
+% SMALLboxInit
+
+if (DL.profile)
+    fprintf('\nStarting Dictionary Learning %s... \n', DL.name);
+end
+
+start=cputime;
+tStart=tic;
+
+%% toolbox configuration
+% test if there is a locally modified version of the config
+% otherwise reads the "default" config file
+run('SMALL_learn_config_local.m');
+
+%%
+%   Dictionary Learning time
+tElapsed=toc(tStart);
+DL.time = cputime - start;
+if (DL.profile)
+    fprintf('\n%s finished task in %2f seconds (cpu time). \n', DL.name, DL.time);
+    fprintf('\n%s finished task in %2f seconds (tic-toc time). \n', DL.name, tElapsed);
+end
+DL.time=tElapsed;
+%   If dictionary is given as a sparse matrix change it to full
+
+DL.D = full(D);
+
+end
+

### 2014-01-08 code/SMALL/SMALL_learn_config_local.m

@@ -1,0 +1,117 @@
+%% Configuration file used in SMALL_learn
+%
+%   Please DO NOT use this file to change the dictionary learning algorithms in SMALLBox
+%   If you want to change the dictionary learning algorithms
+%     create a copy of this file named 'SMALL_learn_config_local.m'
+%
+%   Please refer to the documentation for further information
+
+%   Centre for Digital Music, Queen Mary, University of London.
+%   This file copyright 2009 Ivan Damnjanovic.
+%
+%   This program is free software; you can redistribute it and/or
+%   modify it under the terms of the GNU General Public License as
+%   published by the Free Software Foundation; either version 2 of the
+%   License, or (at your option) any later version.  See the file
+%   COPYING included with this distribution for more information.
+%
+%%
+
+if strcmpi(DL.toolbox,'KSVD')
+    param=DL.param;
+    param.data=Problem.b;
+    
+    D = eval([DL.name,'(param)']);%, ''t'', 5);']);
+elseif strcmpi(DL.toolbox,'KSVDS')
+    param=DL.param;
+    param.data=Problem.b;
+    
+    D = eval([DL.name,'(param, ''t'', 5);']);
+elseif strcmpi(DL.toolbox,'SPAMS')
+    
+    X  = Problem.b;
+    param=DL.param;
+    
+    D = eval([DL.name,'(X, param);']);
+    %   As some versions of SPAMS does not produce unit norm column
+    %   dictionaries, we need to make sure that columns are normalised to
+    %   unit lenght.
+    
+    for i = 1: size(D,2)
+        D(:,i)=D(:,i)/norm(D(:,i));
+    end
+elseif strcmpi(DL.toolbox,'SMALL')
+    
+    X  = Problem.b;
+    param=DL.param;
+    
+    D = eval([DL.name,'(X, param);']);
+    %   we need to make sure that columns are normalised to
+    %   unit lenght.
+    
+    for i = 1: size(D,2)
+        D(:,i)=D(:,i)/norm(D(:,i));
+    end
+    
+elseif strcmpi(DL.toolbox,'TwoStepDL')
+    
+    DL=SMALL_two_step_DL(Problem, DL);
+    
+    %   we need to make sure that columns are normalised to
+    %   unit lenght.
+    
+    for i = 1: size(DL.D,2)
+        DL.D(:,i)=DL.D(:,i)/norm(DL.D(:,i));
+    end
+    D = DL.D;
+    
[…]

### 2014-01-08 code/SMALL/SMALL_solve.m

@@ -1,0 +1,77 @@
+function solver = SMALL_solve(Problem, solver)
+%% SMALL sparse solver caller function
+%
+%   Function gets as input SMALL structure that contains SPARCO problem to
+%   be solved, name of the toolbox and solver, and parameters file for
+%   particular solver.
+%
+%   Outputs are solution, reconstructed signal and time spent
+
+%   Centre for Digital Music, Queen Mary, University of London.
+%   This file copyright 2009 Ivan Damnjanovic.
+%
+%   This program is free software; you can redistribute it and/or
+%   modify it under the terms of the GNU General Public License as
+%   published by the Free Software Foundation; either version 2 of the
+%   License, or (at your option) any later version.  See the file
+%   COPYING included with this distribution for more information.
+%
+%%
+
+% SMALLboxInit
+
+if isa(Problem.A,'float')
+    A = Problem.A;
+    SparseLab_A=Problem.A;
+    m = size(Problem.A,1);      % m is the no. of rows.
+    n = size(Problem.A,2);      % n is the no. of columns.
+else
+    A  = @(x) Problem.A(x,1);  % The operator
+    AT = @(y) Problem.A(y,2);  % and its transpose.
+    SparseLab_A =@(mode, m, n, x, I, dim) SL_A(Problem.A, mode, m, n, x, I, dim);
+    m = Problem.sizeA(1);      % m is the no. of rows.
+    n = Problem.sizeA(2);      % n is the no. of columns.
+    
+end
+% if signal that needs to be represented is different then training set for
+% dictionary learning it should be stored in Problem.b1 matix
+if isfield(Problem, 'b1')
+    b = Problem.b1;
+else
+    b = Problem.b;            % The right-hand-side vector.
+end
+%%
+if (solver.profile)
+    fprintf('\nStarting solver %s... \n', solver.name);
+end
+
+start=cputime;
+tStart=tic;
+
+%% solvers configuration
+% test if there is a locally modified version of the config
+% otherwise reads the "default" config file
+run('SMALL_solve_config.m');
+
+%%
+%   Sparse representation time
+tElapsed=toc(tStart);
+solver.time = cputime - start;
+if (solver.profile)
+    fprintf('Solver %s finished task in %2f seconds (cpu time). \n', solver.name, solver.time);
+    fprintf('Solver %s finished task in %2f seconds (tic-toc time). \n', solver.name, tElapsed);
+end
+solver.time=tElapsed;
[…]

### 2014-01-08 code/SMALL/SMALL_solve_config.m

@@ -1,0 +1,105 @@
+%% Configuration file used in SMALL_solve
+%
+%   Please DO NOT use this file to change the solvers used in SMALLBox
+%   If you want to change the solvers create a copy 
+%     of this file named 'SMALL_learn_config_local.m'
+%
+%   Please refer to the documentation for further information
+
+%   Centre for Digital Music, Queen Mary, University of London.
+%   This file copyright 2009 Ivan Damnjanovic.
+%
+%   This program is free software; you can redistribute it and/or
+%   modify it under the terms of the GNU General Public License as
+%   published by the Free Software Foundation; either version 2 of the
+%   License, or (at your option) any later version.  See the file
+%   COPYING included with this distribution for more information.
+%
+%%
+
+
+if strcmpi(solver.toolbox,'sparselab')
+    y = eval([solver.name,'(SparseLab_A, b, n,',solver.param,');']);
+elseif strcmpi(solver.toolbox,'sparsify')
+    if isa(Problem.A,'float')
+        y = eval([solver.name,'(b, A, n,',solver.param,');']);
+    else
+        y = eval([solver.name,'(b, A, n, ''P_trans'', AT,',solver.param,');']);
+    end
+elseif (strcmpi(solver.toolbox,'spgl1')||strcmpi(solver.toolbox,'gpsr'))
+    y = eval([solver.name,'(b, A,',solver.param,');']);
+elseif (strcmpi(solver.toolbox,'SPAMS'))
+    y = eval([solver.name,'(b, A, solver.param);']);
+elseif (strcmpi(solver.toolbox,'SMALL'))
+    if isa(Problem.A,'float')
+        y = eval([solver.name,'(A, b, n,',solver.param,');']);
+    else
+        y = eval([solver.name,'(A, b, n,',solver.param,',AT);']);
+    end
+elseif (strcmpi(solver.toolbox, 'ompbox'))
+    G=A'*A;
+    epsilon=solver.param.epsilon;
+    maxatoms=solver.param.maxatoms;
+    y = eval([solver.name,'(A, b, G,epsilon,''maxatoms'',maxatoms,''checkdict'',''off'');']);
+    % danieleb: added call to omp functions with fast implementation.
+elseif (strcmpi(solver.toolbox, 'ompbox_fast'))
+    DtX=A'*b;
+    XtX = sum(b.*b);
+    G=A'*A;
+    epsilon=solver.param.epsilon;
[…]

### 2014-01-08 code/SMALL/dico_update.m

@@ -1,0 +1,135 @@
+function [dico, amp] = dico_update(dico, sig, amp, type, flow, rho)
+    
+    %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+    % [dico, amp] = dico_update(dico, sig, amp, type, flow, rho)
+    %
+    % perform one iteration of dictionary update for dictionary learning
+    %
+    % parameters:
+    % - dico: the initial dictionary with atoms as columns
+    % - sig: the training data
+    % - amp: the amplitude coefficients as a sparse matrix
+    % - type: the algorithm can be one of the following
+    %   - ols: fixed step gradient descent, as described in Olshausen &
+    %   Field95
+    %   - opt: optimal step gradient descent, as described in Mailhe et
+    %   al.08
+    %   - MOD: pseudo-inverse of the coefficients, as described in Engan99
+    %   - KSVD: PCA update as described in Aharon06. For fast applications,
+    %   use KSVDbox rather than this code.
+    %   - LGD: large step gradient descent. Equivalent to 'opt' with
+    %          rho=2.
+    % - flow: 'sequential' or 'parallel'. If sequential, the residual is
+    % updated after each atom update. If parallel, the residual is only
+    % updated once the whole dictionary has been computed.
+    % Default: Sequential (sequential usually works better). Not used with
+    % MOD.
+    % - rho: learning rate. If the type is 'ols', it is the descent step of
+    % the gradient (default: 0.1). If the type is 'opt', the
+    % descent step is the optimal step*rho (default: 1, although 2 works
+    % better. See LGD for more details). Not used for MOD and KSVD.
+    %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+    
+    if ~ exist( 'flow', 'var' ) || isempty(flow)
+        flow = 'sequential';
+    end
+    
+    res = sig - dico*amp;
+    nb_pattern = size(dico, 2);
+    
+    % if the type is random, then randomly pick another type
+    switch type
+        case 'rand'
+            x = rand();
+            if x < 1/3
+                type = 'MOD';
+            elseif type < 2/3
+                type = 'opt';
+            else
+                type = 'KSVD';
+            end
+    end
+    
+    % set the learning rate to default if not provided
[…]

### 2014-01-08 code/TestClassSubSpaRan.m

@@ -1,7 +1,16 @@
 function TestClassSubSpaRan
+% add path and compile omp if not already done
+addpath(genpath(pwd));
+A = dir('omp/');
+if length(strfind([A(:).name],'ompmex'))<3 %if the number of files containing 'ompmex' is less than 2 compile
+    cd omp/
+    make
+    cd ..
+end
+
 subSpaRanks = 1:4;
-datasets    = {'coloncancer'};
-datfun = @GetColonCancerDataset;
+datasets    = {'fisher'};
+datfun = @GetFisherIrisDataset;
 par.visu = false;
 for iDat=1:length(datasets)
     fprintf('\nobtaining dataset %s... ',datasets{iDat});

### 2014-01-08 code/TestClassSubSpaRan.m
ignored mexmaci files

@@ -1,9 +1,9 @@
 function TestClassSubSpaRan
 % add path and compile omp if not already done
 addpath(genpath(pwd));
-A = dir('omp/');
+A = dir('SMALL/OMP_Private');
 if length(strfind([A(:).name],'ompmex'))<3 %if the number of files containing 'ompmex' is less than 2 compile
-    cd omp/
+    cd 'SMALL/OMP_Private'
     make
     cd ..
 end

### 2014-01-09 code/Datasets/GetUSPS.m
added usps dataset

@@ -1,0 +1,19 @@
+function [fea,cat] = GetUSPS
+
+load usps_resampled
+
+uniCat = strsplit(num2str(0:9),' ');
+nCat = length(uniCat);
+traTem = mod(find(train_labels==1),10);
+fea = [];
+cat = [];
+for iCat=1:nCat
+    traIdx = traTem==(str2double(uniCat{iCat}));
+    cat = [cat; traTem(traIdx)];
+    fea = [fea, train_patterns(:,traIdx)];
+end
+cat(cat==0) = 10;
+cat = cat-1;
+fea = fea';
+I = @(n) imagesc(reshape(fea(n,:),sqrt(size(fea,2)),sqrt(size(fea,2))));
+clear train_labels train_patterns test_labels test_patterns traTem tesTem

### 2014-01-09 code/Datasets/GetUSPSDataset.m
test subspace dimensions based on dimension of dataset

@@ -1,4 +1,4 @@
-function [fea,cat] = GetUSPS
+function [fea,cat] = GetUSPSDataset
 
 load usps_resampled
 

### 2014-01-09 code/TestClassSubSpaRan.m

@@ -8,22 +8,32 @@ if length(strfind([A(:).name],'ompmex'))<3 %if the number of files containing 'o
     cd ..
 end
 
-subSpaRanks = 1:4;
-datasets    = {'fisher'};
-datfun = @GetFisherIrisDataset;
+% parse dataset functions
+datfun = {@GetFisherIrisDataset,@GetBalanceDataset,...
+    @GetParkinsonsDataset,@GetSonarDataset,@GetUSPSDataset};
+datasets    = cell(length(datfun),1);
+for i=1:length(datfun)
+    s= functions(datfun{i});
+    datasets{i} = regexprep(s.function,'Get(\w+)Dataset','$1');
+end
+
+% parameters
 par.visu = false;
+
 for iDat=1:length(datasets)
     fprintf('\nobtaining dataset %s... ',datasets{iDat});
-    [fea,cat] = datfun();
+    [fea,cat] = datfun{iDat}();                 %get features and categories
     if ~isfloat(fea), fea = double(fea); end     %transform features to float if needed
     fprintf('done\n');
+    feaDim = size(fea,2);
+    subSpaRanks = TrimEnd(ceil(linspace(1,feaDim,5)));
     for iSSR = 1:length(subSpaRanks)
         fprintf('testing transforms using subspaces of rank %d... ',subSpaRanks(iSSR));
-        par.subSpaRan = iSSR;
-        res(iSSR) = TestTransforms(fea,cat,par);
+        par.subSpaRan = subSpaRanks(iSSR);
+        res(iDat,iSSR) = TestTransforms(fea,cat,par);
         fprintf('done\n');
     end
-    mcrs = [res(:).mcr]';
+    mcrs = [res(iDat,:).mcr]';
     figure, PlotMCRs(mcrs(:,2:end));
 end
 
@@ -34,3 +44,11 @@ legend(methodNames);
 xlabel('Sub-space rank');
 ylabel('misclassification ratio');
 grid on
+
+function v = TrimEnd(v)
+% trim the end of a linspace vector if it contains equal values
+ind = find(v==max(v));
+if length(ind) > 1
+    v(ind(2:end)) = [];
+end
+

### 2014-01-09 code/TestClassSubSpaRan.m
improved plot function

@@ -10,7 +10,7 @@ end
 
 % parse dataset functions
 datfun = {@GetFisherIrisDataset,@GetBalanceDataset,...
-    @GetParkinsonsDataset,@GetSonarDataset,@GetUSPSDataset};
+    @GetParkinsonsDataset,@GetSonarDataset};
 datasets    = cell(length(datfun),1);
 for i=1:length(datfun)
     s= functions(datfun{i});
@@ -23,7 +23,7 @@ par.visu = false;
 for iDat=1:length(datasets)
     fprintf('\nobtaining dataset %s... ',datasets{iDat});
     [fea,cat] = datfun{iDat}();                 %get features and categories
-    if ~isfloat(fea), fea = double(fea); end     %transform features to float if needed
+    if ~isfloat(fea), fea = double(fea); end    %transform features to float if needed
     fprintf('done\n');
     feaDim = size(fea,2);
     subSpaRanks = TrimEnd(ceil(linspace(1,feaDim,5)));
@@ -34,15 +34,16 @@ for iDat=1:length(datasets)
         fprintf('done\n');
     end
     mcrs = [res(iDat,:).mcr]';
-    figure, PlotMCRs(mcrs(:,2:end));
+    figure, PlotMCRs(mcrs(:,2:end),subSpaRanks,datasets{iDat});
 end
 
-function PlotMCRs(mcrs)
+function PlotMCRs(mcrs,subSpaRanks,dataset)
 methodNames = {'PCA','SPCA','LDA','IPR'};
-plot(mcrs,'o-','MarkerSize',10)
+plot(subSpaRanks,mcrs,'o-','MarkerSize',10)
 legend(methodNames);
 xlabel('Sub-space rank');
 ylabel('misclassification ratio');
+title(sprintf('Classification of %s dataset',dataset));
 grid on
 
 function v = TrimEnd(v)

### 2014-01-09 code/IPRDis.m
added function to test all main parameters

@@ -8,9 +8,9 @@ end
 %% Parameters and defaults
 if ~exist('par','var') || isempty(par), par = struct; end   %parameters
 
-def.nIte  = 20;     %number of iterations
+def.nIte  = 30;     %number of iterations
 def.eps   = 1e-6;   %tolerance level
-def.alpha = .7;     %mix weight
+def.alpha = .5;     %mix weight
 def.vis   = false;  %visualize results (only for 2 or 3 dimensional datasets)
 def.ver   = false;  %verose output
 
@@ -102,6 +102,8 @@ if iIter<par.nIte
     mus(iIter:end) = mus(iIter);
     srs(iIter:end) = srs(iIter);
 end
+% figure(1), plot(mus,srs)
+% fprintf('\n%f , %f',mus(end),srs(end));
 end
 
 function dic  = ProjectSpec(G,nDim)

### 2014-01-09 code/TestAllParams.m

@@ -1,4 +1,4 @@
-function TestClassSubSpaRan
+function res = TestAllParams
 
 reset(RandStream.getGlobalStream) %reset random stream to reproduce results exactly
 
@@ -23,21 +23,38 @@ end
 % parameters
 par.visu = false;
 
-for iDat=1:length(datasets)
+%pre-allocate output structure
+nDataSets = length(datasets);
+nSubSpaRan = 5;
+nNNN = 3;
+nMus = 2;
+res(nDataSets,nSubSpaRan,nNNN,nMus).mcr = nan;
+res(nDataSets,nSubSpaRan,nNNN,nMus).mets = nan;
+for iDat=1:nDataSets
     fprintf('\nobtaining dataset %s... ',datasets{iDat});
     [fea,cat] = datfun{iDat}();                 %get features and categories
     if ~isfloat(fea), fea = double(fea); end    %transform features to float if needed
     fprintf('done\n');
     feaDim = size(fea,2);
     subSpaRanks = TrimEnd(ceil(linspace(1,feaDim,5)));
-    for iSSR = 1:length(subSpaRanks)
+    for iSSR = 1:nSubSpaRan
         fprintf('testing transforms using subspaces of rank %d... ',subSpaRanks(iSSR));
         par.subSpaRan = subSpaRanks(iSSR);
-        res(iDat,iSSR) = TestTransforms(fea,cat,par);
+        nns = [1,5,floor(size(fea,1)/length(unique(cat)))]; %number of knn neighbours
+        for inn=1:nNNN
+            par.nKNNs = nns(inn);
+            mus = [0,1];
+            for imu=1:nMus
+                par.mu = mus(imu);
+                tmp = TestTransforms(fea,cat,par);
+                res(iDat,iSSR,inn,imu).mcr = tmp.mcr;
+                res(iDat,iSSR,inn,imu).mets = tmp.mets;
+            end
+        end
         fprintf('done\n');
     end
-    mcrs = [res(iDat,:).mcr]';
-    figure, PlotMCRs(mcrs(:,2:end),subSpaRanks,datasets{iDat});
+%     mcrs = [res(iDat,:).mcr]';
+%     figure, PlotMCRs(mcrs(:,2:end),subSpaRanks,datasets{iDat});
 end
 
 function PlotMCRs(mcrs,subSpaRanks,dataset)

### 2014-01-09 code/TestTransforms.m

@@ -1,7 +1,6 @@
 function res = TestTransforms(fea,cat,par)
 
 addpath(genpath(pwd)); %add path to matlab search path
-reset(RandStream.getGlobalStream) %reset random stream to reproduce results exactly
 
 %% Parameters and defaults
 if ~exist('par','var') || isempty(par), par = struct; end

### 2014-01-09 code/TestTransforms.m
updated parallel options

@@ -5,6 +5,8 @@ addpath(genpath(pwd)); %add path to matlab search path
 %% Parameters and defaults
 if ~exist('par','var') || isempty(par), par = struct; end
 
+options = statset('UseParallel',true); %set option to use parallel computing
+
 def.nAto = 2*size(fea,2); %number of atoms (twice overcomplete dictionary)
 def.subSpaRan = 1; %subspace range
 def.perActAto = 50; %percentage of active atoms
@@ -40,7 +42,7 @@ tesFea = (tesFea-repmat(feaMea,nObs-nTraObs,1))./repmat(feaStd,nObs-nTraObs,1); 
 cvPar = cvpartition([traCat;tesCat],'kfold',5);                 %create 5fold partition
 
 conMatFun = @(traFea,traCat,tesFea,tesCat)TransformAndClassify(traFea,traCat,tesFea,tesCat,par);
-conMat = crossval(conMatFun,[traFea;tesFea],nominal([traCat;tesCat]),'partition',cvPar); %compute crossvalidation results on original features
+conMat = crossval(conMatFun,[traFea;tesFea],nominal([traCat;tesCat]),'partition',cvPar,'options',options); %compute crossvalidation results on original features
 
 nCat = length(unique(nominal([traCat;tesCat]))); %number of categories
 conMat = reshape(sum(conMat),nCat,nCat); %confusion matrix
@@ -49,28 +51,28 @@ mcr = sum(sum(conMat-diag(diag(conMat))))/sum(sum(conMat)); %misclassification r
 %% Test PCA feature transform
 [pcaTraFea, pcaTesFea] = PCAFeaturesTransform(traFea,tesFea,par); %compute PCA feature transform
 conMatFun = @(traFea,traCat,tesFea,tesCat)PCATransformAndClassify(traFea,traCat,tesFea,tesCat,par);
-pcaConMat = crossval(conMatFun,[traFea;tesFea],nominal([traCat;tesCat]),'partition',cvPar);
+pcaConMat = crossval(conMatFun,[traFea;tesFea],nominal([traCat;tesCat]),'partition',cvPar,'options',options);
 pcaConMat = reshape(sum(pcaConMat),nCat,nCat);
[…]

### 2014-01-10 code/Datasets/GetToyExampleDataset.m
added toy example dataset

@@ -1,0 +1,25 @@
+function [fea,cat] = GetToyExampleDataset
+
+clear, clc, close all
+reset(RandStream.getGlobalStream);    %set random number generator to reproduce results
+
+nData = 1000;						%number of data for each class
+theta = [0, pi/4, pi/2];            %angles
+nThetas = length(theta);			%number of angles
+fea = zeros(nThetas*nData,2);
+cat = zeros(nThetas*nData,1);
+for iTheta=1:nThetas
+    fea((iTheta-1)*nData+1:iTheta*nData,:) = randomsamp(nData,theta(iTheta));
+    cat((iTheta-1)*nData+1:iTheta*nData) = iTheta*ones(nData,1);
+end
+% figure
+% gscatter(fea(:,1),fea(:,2),cat,'krb','.',10);
+end
+
+function X = randomsamp(nData,theta)
+eps = 1e-1;
+R = [cos(theta), -sin(theta); sin(theta), cos(theta)]; %rotation matrix
+v = [linspace(-1,1,nData); zeros(1,nData)];
+X = R*v;
+X = (X+eps*randn(size(X)))';
+end

### 2014-01-10 code/TestClassSubSpaRan.m

@@ -12,7 +12,7 @@ if length(strfind([A(:).name],'ompmex'))<3 %if the number of files containing 'o
 end
 
 % parse dataset functions
-datfun = {@GetFisherIrisDataset,@GetBalanceDataset,...
+datfun = {@GetToyExampleDataset,@GetFisherIrisDataset,@GetBalanceDataset,...
     @GetParkinsonsDataset,@GetSonarDataset};
 datasets    = cell(length(datfun),1);
 for i=1:length(datfun)

### 2014-01-10 code/TestTransforms.m

@@ -3,6 +3,13 @@ function res = TestTransforms(fea,cat,par)
 addpath(genpath(pwd)); %add path to matlab search path
 
 %% Parameters and defaults
+if ~exist('fea','var') || isempty(fea) || ~exist('cat','var') || isempty(cat)
+    [fea,cat] = GetUSPSDataset([1,3,8]);
+    if isnumeric(cat), cat = cellstr(num2str(cat)); end
+%     nDim = 4; %data has 4 dimensions
+%     fea = fea(:,1:nDim);
+end
+
 if ~exist('par','var') || isempty(par), par = struct; end
 
 options = statset('UseParallel',true); %set option to use parallel computing
@@ -10,16 +17,12 @@ options = statset('UseParallel',true); %set option to use parallel computing
 def.nAto = 2*size(fea,2); %number of atoms (twice overcomplete dictionary)
 def.subSpaRan = 1; %subspace range
 def.perActAto = 50; %percentage of active atoms
-def.nKNNs = 5; %number of nearest neighbours for classification
-def.visu = false;
+def.nKNNs = 1; %number of nearest neighbours for classification
+def.visu = true;
 
 par = setdefaultoptions(par,def);
 
-if ~exist('fea','var') || isempty(fea) || ~exist('cat','var') || isempty(cat)
-    [fea,cat] = GetFisherIrisDataset; %get data from iris dataset
-    nDim = 4; %data has 4 dimensions
-    fea = fea(:,1:nDim);
-end
+
 
 % Divide observations in training and test sets (for visualization)
 [nObs,nDim] = size(fea);
@@ -90,32 +93,36 @@ if par.visu
     for iMet=1:nMet
         switch mets{iMet}
             case 'none'
-                fea = traFea;
+                traFea = traFea;
                 mat = conMat;
                 err = mcr;
             case 'PCA'
-                fea = [pcaTraFea zeros(size(pcaTraFea,1),1)];
+                traFea = [pcaTraFea zeros(size(pcaTraFea,1),1)];
                 tesFea = [pcaTesFea zeros(size(pcaTesFea,1),1)];
                 mat = pcaConMat;
                 err = pcaMcr;
             case 'SPCA'
-                fea = [spcaTraFea zeros(size(spcaTraFea,1),1)];
+                traFea = [spcaTraFea zeros(size(spcaTraFea,1),1)];
[…]

### 2014-01-10 code/Datasets/GetUSPSDataset.m
updated usps experiment

@@ -1,19 +1,34 @@
-function [fea,cat] = GetUSPSDataset
+function [fea,cat] = GetUSPSDataset(digits)
 
 load usps_resampled
 
 uniCat = strsplit(num2str(0:9),' ');
 nCat = length(uniCat);
 traTem = mod(find(train_labels==1),10);
-fea = [];
-cat = [];
+tmpFea = [];
+tmpCat = [];
 for iCat=1:nCat
     traIdx = traTem==(str2double(uniCat{iCat}));
-    cat = [cat; traTem(traIdx)];
-    fea = [fea, train_patterns(:,traIdx)];
+    tmpCat = [tmpCat; traTem(traIdx)];
+    tmpFea = [tmpFea, train_patterns(:,traIdx)];
 end
-cat(cat==0) = 10;
-cat = cat-1;
-fea = fea';
+tmpCat(tmpCat==0) = 10;
+tmpCat = tmpCat-1;
+tmpFea = tmpFea';
+
+tmpCat = cellstr(num2str(tmpCat));
+cat = [];
+fea = [];
+if exist('digits','var')
+    for iDig = 1:length(digits)
+        idx = strcmp(num2str(digits(iDig)),tmpCat);
+        cat = [cat; tmpCat(idx)];   
+        fea = [fea; tmpFea(idx,:)];
+    end
+else
+    fea = tmpFea;
+    cat = tmpCat;
+end
+
 I = @(n) imagesc(reshape(fea(n,:),sqrt(size(fea,2)),sqrt(size(fea,2))));
 clear train_labels train_patterns test_labels test_patterns traTem tesTem

### 2014-01-10 code/IPRLearnDisSub.m

@@ -13,7 +13,7 @@ nDim = size(traFea,2);                  %dimension of features
 toolbox   = 'SMALL_incoherentDL';		%dictionary learning toolbox
 
 def.subSpaRan = 50;                     %rank of incoherent subspaces
-def.nAto = nCat*nDim;                   %number of atoms
+def.nAto = 2*nDim;                      %number of atoms
 def.nIte = 1;                           %number of DL iterations
 
 def.mu = sqrt((par.nAto-nDim)/(nDim*(par.nAto-1))); %mutual coherence
@@ -113,7 +113,7 @@ gscatter(traFea(:,1), traFea(:,2), traCat,'rgb','osd');
 xlabel('Sepal length');
 ylabel('Sepal width');
 
-newTraFea = IPRFeaturesTransform(traFea,tesFea,traCat);
+[traSubSpa,newTraFea,newTesFea] = IPRLearnDisSub(traFea,tesFea,traCat);
 size(newTraFea,2);
 figure
 if size(newTraFea,2)<2

### 2014-01-10 code/TestClassSubSpaRan.m

@@ -12,8 +12,7 @@ if length(strfind([A(:).name],'ompmex'))<3 %if the number of files containing 'o
 end
 
 % parse dataset functions
-datfun = {@GetToyExampleDataset,@GetFisherIrisDataset,@GetBalanceDataset,...
-    @GetParkinsonsDataset,@GetSonarDataset};
+datfun = {@GetUSPSDataset};
 datasets    = cell(length(datfun),1);
 for i=1:length(datfun)
     s= functions(datfun{i});
@@ -25,7 +24,7 @@ par.visu = false;
 
 for iDat=1:length(datasets)
     fprintf('\nobtaining dataset %s... ',datasets{iDat});
-    [fea,cat] = datfun{iDat}();                 %get features and categories
+    [fea,cat] = datfun{iDat}([1,3,8]);                 %get features and categories
     if ~isfloat(fea), fea = double(fea); end    %transform features to float if needed
     fprintf('done\n');
     feaDim = size(fea,2);

### 2014-01-10 code/TestTransforms.m

@@ -17,7 +17,7 @@ options = statset('UseParallel',true); %set option to use parallel computing
 def.nAto = 2*size(fea,2); %number of atoms (twice overcomplete dictionary)
 def.subSpaRan = 1; %subspace range
 def.perActAto = 50; %percentage of active atoms
-def.nKNNs = 1; %number of nearest neighbours for classification
+def.nKNNs = 5; %number of nearest neighbours for classification
 def.visu = true;
 
 par = setdefaultoptions(par,def);

### 2014-01-10 code/LDAFeaturesTransform.m
changed graph and updated lda to always project data onto nUniCat-1 dimensional subspace

@@ -9,8 +9,10 @@ St = Sb + Sw;                               %Total scatter matrix
 J = @(W) trace(pinv(W'*St*W)*(W'*Sb*W));    %as a function of projection matrix
 
 [D,X] = eig(pinv(St)*Sb);                   %compute evd
-idx = diag(real(X))>1e-8;                         %find non-zero eigenvalues
-W = real(D(:, idx));                              %select corresponding eigenvectors
+idx = diag(real(X))>1e-8;                   %find non-zero eigenvalues
+W = real(D(:, idx));                        %select corresponding eigenvectors
+W = real(D(:,1:length(unique(traCat))-1));
+% assert(size(W,2)==(length(unique(traCat))-1));
 
 newTraFea = (W*pinv(W)*traFea')';           %project data onto subspace spanned by eigenvectors
 newTesFea = (W*pinv(W)*tesFea')';

### 2014-01-10 code/TestClassSubSpaRan.m

@@ -12,7 +12,8 @@ if length(strfind([A(:).name],'ompmex'))<3 %if the number of files containing 'o
 end
 
 % parse dataset functions
-datfun = {@GetUSPSDataset};
+datfun = {@GetToyExampleDataset,@GetFisherIrisDataset,@GetBalanceDataset,...
+    @GetParkinsonsDataset,@GetSonarDataset};
 datasets    = cell(length(datfun),1);
 for i=1:length(datfun)
     s= functions(datfun{i});
@@ -24,7 +25,7 @@ par.visu = false;
 
 for iDat=1:length(datasets)
     fprintf('\nobtaining dataset %s... ',datasets{iDat});
-    [fea,cat] = datfun{iDat}([1,3,8]);                 %get features and categories
+    [fea,cat] = datfun{iDat}();                 %get features and categories
     if ~isfloat(fea), fea = double(fea); end    %transform features to float if needed
     fprintf('done\n');
     feaDim = size(fea,2);
@@ -41,12 +42,12 @@ end
 
 function PlotMCRs(mcrs,subSpaRanks,dataset)
 methodNames = {'PCA','SPCA','LDA','IPR'};
-plot(subSpaRanks,mcrs,'o-','MarkerSize',10)
+plot(subSpaRanks,mcrs,'o--','MarkerSize',10)
 legend(methodNames);
 xlabel('Sub-space rank');
 ylabel('misclassification ratio');
 title(sprintf('Classification of %s dataset',dataset));
-grid on
+set(gca,'YGrid','on','XGrid','off','XTick',subSpaRanks);
 
 function v = TrimEnd(v)
 % trim the end of a linspace vector if it contains equal values

### 2014-01-10 code/TestTransforms.m

@@ -4,10 +4,10 @@ addpath(genpath(pwd)); %add path to matlab search path
 
 %% Parameters and defaults
 if ~exist('fea','var') || isempty(fea) || ~exist('cat','var') || isempty(cat)
-    [fea,cat] = GetUSPSDataset([1,3,8]);
+    [fea,cat] = GetFisherIrisDataset;
     if isnumeric(cat), cat = cellstr(num2str(cat)); end
-%     nDim = 4; %data has 4 dimensions
-%     fea = fea(:,1:nDim);
+    nDim = 3; %data has 4 dimensions
+    fea = fea(:,1:nDim);
 end
 
 if ~exist('par','var') || isempty(par), par = struct; end
@@ -84,7 +84,7 @@ iprMcr = sum(sum(iprConMat-diag(diag(iprConMat))))/sum(sum(iprConMat));
 % J1 = @(fea,cat) trace(pinv(ClassificationDiscriminant.fit(fea,cat).Sigma)*ClassificationDiscriminant.fit(fea,cat).BetweenSigma);
 % J2 = @(fea,cat) trace(ClassificationDiscriminant.fit(fea,cat).BetweenSigma)/trace(ClassificationDiscriminant.fit(fea,cat).Sigma);
 % J3 = @(fea,cat) log(abs(det(ClassificationDiscriminant.fit(fea,cat).BetweenSigma))) - ClassificationDiscriminant.fit(fea,cat).LogDetSigma;
-mets = {'none','PCA','SPCA','LDA','IPR'};
+mets = {'none','IPR'};
 if par.visu
     nMet = length(mets);
     sqrNMet = ceil(sqrt(nMet));
@@ -97,13 +97,17 @@ if par.visu
                 mat = conMat;
                 err = mcr;
             case 'PCA'
-                traFea = [pcaTraFea zeros(size(pcaTraFea,1),1)];
-                tesFea = [pcaTesFea zeros(size(pcaTesFea,1),1)];
+                if size(pcaTraFea,2)~=3
+                    traFea = [pcaTraFea zeros(size(pcaTraFea,3-size(pcaTraFea,2)),1)];
+                    tesFea = [pcaTesFea zeros(size(pcaTesFea,3-size(pcaTraFea,2)),1)];
+                end
                 mat = pcaConMat;
                 err = pcaMcr;
             case 'SPCA'
-                traFea = [spcaTraFea zeros(size(spcaTraFea,1),1)];
[…]

### 2014-01-13 manuscript.tex
added image and code block
added pseudo code generation
added function to load and display fishers dataset
added functions for reproducing mlsp figure
removed org version because I have no time, added files and started expanding mlsp paper from background section

--- 2013-12-13 manuscript.tex
@@
-% Created 2013-12-13 Fri 11:53
-\documentclass[]{article}
-\usepackage[utf8]{inputenc}
+\documentclass{article}
+
+\title{Learning Incoherent Subspaces: Classification via Incoherent Dictionary Learning.}
+
+\author{Daniele Barchiesi and Mark D. Plumbley}
+%\affiliation{Centre for Digital Music\\
+%	Queen Mary University of London\\
+%	Mile End Road, London E1 4NS, UK}
+
+%%%%%%%%%%%%%%%%%%%%%%PACKAGES%%%%%%%%%%%%%%%%%%%%%
+\usepackage{amsmath,graphicx}
+\usepackage[english]{babel}
+\usepackage[applemac]{inputenc}
 \usepackage[T1]{fontenc}
-\usepackage{fixltx2e}
-\usepackage{graphicx}
-\usepackage{longtable}
-\usepackage{float}
-\usepackage{wrapfig}
-\usepackage{rotating}
-\usepackage[normalem]{ulem}
-\usepackage{amsmath}
-\usepackage{textcomp}
-\usepackage{marvosym}
-\usepackage{wasysym}
-\usepackage{amssymb}
-\usepackage{hyperref}
-\tolerance=1000
-\author{Daniele Barchiesi}
-\date{}
-\title{Dictionary learning for classification: learning incoherent subspaces}
-\hypersetup{
-  pdfkeywords={},
-  pdfsubject={},
-  pdfcreator={Emacs 24.3.1 (Org mode 8.2.4)}}
+\usepackage{epsfig}
+\usepackage{amsfonts,amssymb,amsmath,amsthm,bm}
+\usepackage[boxruled,linesnumbered]{algorithm2e}
+\usepackage{subfig,graphicx}
+\usepackage{setspace}
+\usepackage{enumitem}
+
+%%%%%%%%%%%%%%%%%%%%%DEFINITIONS%%%%%%%%%%%%%%%%%%
+\input{definitions.tex}
+
+\def \nComponents{L} 	%number of significant components of ICA feature transform
+\def \fea{\Vector{x}} 	%vector of features
+\def \Feas{\Matrix{X}} 	%matrix containing set of training features
+\def \iFea{n}	 		%component index of features vector
+\def \nDim{N} 			%dimensionality of features vector
+\def \newFea{\Vector{y}} 	%features vector after feature transform
+\def \NewFeas{\Matrix{Y}}%matrix of training features after feature transform
+\def \Dic{\Matrix{\Phi}} %dictionary learned from features
+\def \nAto{K} 			%number of atoms in the dictionary
+\def \iAto{k} 			%atom index
+\def \atom{\Vector{\phi}}%atom in a dictionary
+\def \nFea{M}			%number of signals or features
+\def \iFea{m}			%signal index
+\def \Coeff{\Matrix{A}} 	%matrix of sparse approximation coefficients
+\def \coherence{\mu} 		%mutual coherence
+\def \coeff{\Vector{\alpha}}%vector of sparse approximation coefficients
+\def \nActiveAtoms{S} 	%number of active atoms
+\def \orthmat{\Matrix{W}} 	%orthonormal matrix
+\def \cat{c} 			%category
+\def \cats{\Vector{c}}	%vector of categories of training signals
+\def \Cat{\mathcal{C}} 	%set of possible categories
+\def \uniCat{C} 			%element of the set of possible categories
+\def \iCat{p}  		%index of the elements in the set of possible categories
+\def \nCat{P} 		%number of elements in the set of possible categories
+\def \definition{\overset{\Label{def}}{=}}	%definitions
+\def \Gram{\Matrix{G}} 						%Gram matrix
+\def \gram{g} 								%element of the Gram matrix
+\def \admissibleDictionary{\Function{D}} 	%set of admissible dictionaries
+\def \ambient{\Set{R}} 						%ambient
+\def \ipr{\Acronym{ipr}} 					%Iterative projections and rotations
+\def \ip{\Acronym{ip}}						%Iterative projections
+\def \nDimSub{Q}
+\def \Spa{\Matrix{\Psi}}
+
 \begin{document}
+%\ninept
+%
 
 \maketitle
+%
+\begin{abstract}
 
-\section{Abstract}
-\label{sec-1}
-\section{Introduction}
-\label{sec-2}
-\section{Background}
-\label{sec-3}
-\section{Incoherent Dictionary Learning}
-\label{sec-4}
-\section{Learning Incoherent Subspaces}
-\label{sec-5}
-\section{Numerical Experiments}
-\label{sec-6}
-\section{Discussion}
-\label{sec-7}
-\section{Conclusions}
-\label{sec-8}
-% Emacs 24.3.1 (Org mode 8.2.4)
+\end{abstract}
+%
+%\begin{keywords}
+%Feature transforms, sparse approximation, dictionary learning, supervised classification.
+%\end{keywords}
+%
+\section{Introduction: Classification And Feature Transforms}\label{sec:intro}
+Supervised classification is one of the classic problems in machine learning where a system is designed to discriminate the category of an observed signal, having previously observed representative examples from the considered classes \cite{Duda1973Pa}.
+
+Typically, a classification algorithm consists of a training phase where  class-specific models are learned from labelled samples, followed by a testing phase where unlabelled data are classified by comparison with the learned models.
+Both training and testing comprise various stages.
+Firstly, we observe a signal that measures a process of interest, such as the recording of a sound or image, or a log of the temperatures in a particular geographic area.
+Then, a set of features are extracted from the raw signals using signal processing techniques.
+This step is performed in order to reduce the dimensionality of the data and provide a new signal that allows generalisation among examples of the same class, while retaining enough information to discriminate between different classes.
+
+Following the features extraction step, a feature transform can be employed to further reduce the dimensionality of the data and to enhance discrimination between classes.
+Thus classification benefits from feature transforms especially when features are not separable, that is, when it is not possible to optimise a simple function that maps features belonging to signals of a given class to the corresponding category.
+A further dimensionalty reduction may be performed when dealing with high dimensional signals (such as audio or high resolution images) by fitting the parameters of global statistical distributions with features learned on portions of the signal.
+Models learned on different classes are finally compared using a distance metric to the model learned form an unlabelled signal, which is typically assigned to the nearest class.
+
+\subsection{Traditional Algorithms For Feature Transform}\label{sec:aft}
+Two of the main feature transform techniques include principal component analysis (\Acronym{pca}) \cite{Pearson1901On} and Fisher's linear discriminant analysis (\Acronym{lda}) \cite{Duda1973Pa}.
+
+\subsubsection{\Acronym{pca}}
+Let $\curlyb{\fea_{\iFea} \in \real^{\nDim}}_{\iFea=1}^{\nFea}$ be a set of vectors containing features extracted from $\nFea$ training signals.
+The goal of \Acronym{pca} is to learn an orthonormal set of basis functions $\curlyb{\atom_{\iAto} \in \ambient^{\nDim}}_{\iAto=1}^{\nDim}$ such that $ \norm{\atom_{\iAtom}}{2}=1$ and $\inner{\atom_{i}}{\atom_{j}}=0 \; \forall i\neq j$ that are placed along the columns of a so-called \emph{dictionary} $\Dic\in\real^{\nDim\times\nDim}$.
+The bases are optimised from the data to identify their principal components, that is, the sub-spaces that retain the maximum variance of the features.
+First, the eigenvalue decomposition of the outer product $\Matrix{X}\Transpose{\Matrix{X}} = \Matrix{Q}\Matrix{\Lambda}\Transpose{\Matrix{Q}}$ is computed.
+Then, the largest $\nComponents$ eigenvalues are selected from the matrix $\Matrix{\Lambda}$, and the corresponding eigenvectors are
+
+Let $\Spa\definition\Dic_{1:\nComponents}\in\ambient^{\nDim\times\nComponents}$ denote the sub-dictionary constructed from the $\nComponents<\nDim$ principal components of the dataset.
+A new set of transformed features $\newFea_{\Acronym{pca}} = \Spa\Transpose{\Spa}\fea$ is computed by projecting the data onto the sub-space spanned by the columns of $\Dic_{1:\nComponents}$ (that is, onto the $\nComponents$-dimensional principal sub-space).
+This operation reduces the dimensionality of the features by projecting them onto a linear subspace embedded in $\real^{\nDim}$.
+It is an unsupervised technique that does not exploit knowledge about the classes associated with the training set, but implicitly relies in the assumption that the principal component directions encode relevant differences between classes.
+
+\subsubsection{\Acronym{lda}}
+In contrast, \Acronym{lda} is a supervised method for feature transform whose objective is to explicitly maximise the separability of classes in the transformed domain.
+The within-classes scatter matrix is defined to measure how features belonging to the same class are clustered around their mean vector (the vector obtained by averaging all observations associated with a given category), and the  between-classes scatter matrix is defined to measure the distances between the mean vectors.
+A square linear transform $\Matrix{M} \in \real^{\nDim\times\nDim}$ is optimised to obtain a new set of features $\newFea_{\Acronym{lda}}=\Matrix{M}\fea$ according to an objective that promotes features belonging to the same class to be near each other (according to a Euclidean distance), and far apart from features belonging to different classes.
+\Acronym{lda} explicitly seeks to enhance the discriminative power of features, but does not perform dimensionality reduction.
+
+\subsection{Supervised dimensionality reduction}
+\subsubsection{Supervised PCA}
+Related works that extend \Acronym{pca} include the supervised \Acronym{pca} (\Acronym{s-pca}) proposed by Barshan et al.
+\cite{Barshan2011Su}.
+\Acronym{s-pca} is based on the theory of reproducing kernel Hilbert spaces (\Acronym{rkhs}) (that are spaces of functions which satisfy certain properties and map elements from an arbitrary set to the set of complex numbers) \cite{Aronszajn:1950}, and on the so-called Hilbert-Schmidt independence criterion (\Acronym{hsic})\cite{gretton2005measuring}.
+The \Acronym{hsic} is used to estimates the statistical dependence of two random variables based on the fact that this quantity is related to the correlation of functions belonging to their respective \Acronym{rkhs}.
+While \Acronym{hsic} is defined in terms of the probability density function of the two random variables,  empirical estimates of \Acronym{hsic} can be obtained from finite sequences of their realisations.
+The empirical \Acronym{hsic} can be used in turn to construct an objective function that maximises the dependence between the two variables.
+Hence, this strategy is adopted within the context of classification to maximise the statistical dependence between a transformed feature $\newFea_{\Acronym{s-pca}}$ and its corresponding category $\cat$.
+
+In practice, \Acronym{s-pca} differs from \Acronym{pca} in that it calculates the eigenvalue decomposition of a matrix $\Matrix{Q}$ defined as follows:
+\begin{equation}
+\Matrix{Q} \definition \Matrix{X}\Matrix{H}\Matrix{L}\Matrix{H}\Transpose{\Matrix{X}}
+\end{equation}
+were $\Matrix{H} \definition \Matrix{I} - \\Vector{e}\Transpose{\Vector{e}}$ is a \emph{centring} matrix and $\Matrix{L} \definition \Vector{\cat}\Transpose{\Vector{\cat}}$ is the kernel matrix of the class variable that is constructed by computing the outer product of the vectors resulting from assigning different numerical values to each category.
+
+\subsection{Incoherent subspace learning and other related work}
+
+and methods for manifold learning reviewed by Van Der Maaten et al.
+\cite{Van-Der-Maaten2009Di}.
+Finally, the sparse sub-space clustering technique developed by Elhamifar and Vidal \cite{Elhamifar2013Sp} applies concepts and algorithm from the field of sparse approximation to tackle unsupervised clustering problems.
+
+\subsection{Main Contributions and Paper organisation}
+The method proposed in this paper is aimed at learning discriminative sub-spaces that allow dimensionality reduction, while at the same time enhancing the separability between classes.
+It is derived from our previous work on learning incoherent dictionaries for sparse approximation.
+
+Standard methods for feature transform will be briefly reviewed in Section \ref{sec:aft}, as their limitations lead to the main motivation for this work.
+To provide the context of the proposed method, the incoherent dictionary learning problem will be introduced in Section \ref{sec:idl}, while Section \ref{sec:lis} will contain the main contribution of this paper consisting in learning incoherent subspaces for classification.
+Numerical experiments are presented in Section \ref{sec:ne}, and conclusions are drawn in Section \ref{sec:end}.
+
+\section{Incoherent Dictionary Learning}\label{sec:idl}
+A sparse approximation of a signal $\fea\in\real^{\nDim}$ is a linear combination of $\nAto\geq\nDim$ basis functions $\curlyb{\atom_{\iAto}\in\real^{\nDim}}_{\iAto=1}^{\nAto}$ called \emph{atoms} described by:
+\begin{equation}
+	\fea \approx \approximant{\fea} = \sum_{\iAto=1}^{\nAto} \alpha_{\iAto}\atom_{\iAto}
+\end{equation}
+where the vector of coefficients $\coeff$ contains a \emph{small} number of non-zero components, corresponding to a small number of atoms actively contributing to the approximation $\approximant{\fea}$.
+Given a signal $\fea$ and a dictionary, various algorithms have been proposed to find a sparse approximation that minimises the residual error $\norm{\fea-\approximant{\fea}}{2}$\cite{Elad2010Sp}.
+
+Dictionary learning aims at optimising a dictionary $\Dic$ for sparse approximation given a set of training data.
+It is an unsupervised technique that can be thought as being a generalisation of \Acronym{pca}, as both methods learn linear subspaces that minimise the approximation error of the signals.
+Dictionary learning, however, is generally more flexible than \Acronym{pca} because it can be employed to learn more general non-orthogonal over-complete dictionaries \cite{Rubinstein2010Di}.
+
+\subsection{The incoherent dictionary learning problem}
+Dictionaries for sparse approximation have important intrinsic properties that describe the relations between their atoms, like the mutual coherence $\mu(\Dic)=\underset{i\neq j}{\max}{\inner{\atom_{i}}{\atom_{j}}}$ that is defined as the maximum inner product between any two different atoms.
+The goal of incoherent dictionary learning is to learn atoms that are well adapted to sparsely approximate a set of training signals, and that are at the same time mutually incoherent \cite{Barchiesi2013Le}.
+
+Given a set of $\nFea$ training signals contained in the columns of the matrix $\Feas \in \real^{\nDim\times\nFea}$ and a matrix $\Coeff\in\real^{\nAto\times\nFea}$ indicating the sparse approximation coefficients, the incoherent dictionary learning problem can be expressed as:
+
+\begin{align}\label{eq:iprcost}
+\optimal{\Dic} = \MinimiseST{\Dic}{\norm{\Feas-\Dic\Coeff}{\F}}{\coherence(\Dic) \leq \coherence_{0} \nonumber \\
+			&\norm{\coeff_{\iFea}}{0}\leq \nActiveAtoms \quad \forall \iFea}
+\end{align}
+where $\coherence_{0}$ is a fixed mutual coherence constraint, the $\ell_{0}$ pseudo-norm $\norm{\cdot}{0}$ counts the number of non-zero components of its argument and $\nActiveAtoms$ is a fixed number of active atoms.
+Algorithms for (incoherent) dictionary learning generally follow an alternate optimisation heuristic, iteratively updating $\Dic$ and $\Coeff$ until a stopping criterion is met.
+In the case of the iterative projections and rotations algorithm (\Acronym{ipr}) algorithm \cite{Barchiesi2013Le}, a dictionary de-correlation step is added after updating the dictionary in order to satisfy the mutual coherence constraint.
+
+Given $\Feas$, fixed $\coherence_{0}$, $\nActiveAtoms$ and a stopping criterion (such as a maximum number of iterations), the optimisation of \eqref{eq:iprcost} is tackled by iteratively performing the following steps:
+\begin{itemize}
+	\item\emph{Sparse coding}: fix $\Dic$ and compute the matrix $\Coeff$ using a suitable sparse approximation method.
+	\item\emph{Dictionary update}: fix $\Coeff$ and update $\Dic$ using a suitable method for dictionary learning.
+	\item\emph{Dictionary de-correlation}: given $\Feas$, $\Dic$ and $\Coeff$ update the dictionary $\Dic$ to reduce its mutual coherence under the level $\coherence_{0}$.
+\end{itemize}
+
+\subsection{The iterative projections and rotations algorithm}\label{sec:ipr}
+The \Acronym{ipr} algorithm has been proposed in order to solve the dictionary de-correlation step, while ensuring that the updated dictionary provides a sparse approximation with low residual norm, as indicated by the objective function \eqref{eq:iprcost} \cite{Barchiesi2013Le}.
+
+The \Acronym{ipr} algorithm requires the calculation of the Gram matrix $\Gram=\Transpose{\Dic}\Dic$ which contains the inner products between any two atoms in the dictionary.
+$\Gram$ is iteratively projected onto two constraint sets, namely the structural constraint set $\stcset$ and the spectral constraint set $\spcset$.
+The former is the set of symmetric square matrices with unit diagonal values and off-diagonal values with magnitude smaller or equal than $\coherence_{0}$:
+\small
+\begin{equation*}
+	\stcset \definition \curlyb{\stcmat \in \ambient^{\nAtoms \times \nAtoms} : \stcmat = \Transpose{\stcmat}, \stcel_{i,i}=1,\max_{i > j}|\stcel_{i,j}|\leq \coherence_{0}}.
+\end{equation*}
+\normalsize
+The latter is the set of symmetric positive semidefinite square matrices with rank smaller than or equal to $\nDimensions$:
+\begin{equation*}
+\spcset \definition \curlyb{ \spcmat \in \ambient^{\nAtoms \times \nAtoms} : \spcmat = \Transpose{\spcmat}, \operatorname{eig}(\spcmat)\geq \Vector{0}, \operatorname{rank}(\spcmat)\leq \nDimensions}
+\end{equation*}  
+where the operator $\operatorname{eig}(\cdot)$ returns the vector of eigenvalues of its argument.
+
+Starting from the Gram matrix of an initial dictionary $\Dictionary$, the \Acronym{ipr} method iteratively performs the following operations.
+\begin{itemize}
+\item \emph{Projection onto the structural constraint set}. The projection $\stcmat = \Projection_{\stcset}(\Gram)$ can be obtained by:
+\begin{enumerate}
+\item setting $\stcel_{i,i} = 1$,
+\item limiting the off-diagonal elements so that, for $i \neq j$, 
+\small
+\begin{equation}\label{eq:pscs}
+\stcel_{i,j} = \operatorname{Limit}({\gram}_{i,j},\coherence_{0}) = \left\{ \begin{array}{rl}
+	\gram_{i,j}  & \text{if} \quad |\gram_{i,j}|\leq \coherence_{0} \\
+	\operatorname{sgn}(\gram_{i,j})\coherence_{0}  & \text{if} \quad |\gram_{i,j}| > \coherence_{0}
+\end{array} \right.
+\end{equation}
+\normalsize
+\end{enumerate}
+\item \emph{Projection onto the spectral constraint set and factorization}. The projection $\spcmat = \Projection_{\spcset}(\Gram)$ and subsequent factorisation are obtained by:
+\begin{enumerate}
+\item calculating the eigenvalue decomposition (\Acronym{evd}) $\Gram = \eigvecmat \eigvalmat \Transpose{\eigvecmat}$,
+\item thresholding the eigenvalues by keeping only the $\nDimensions$ largest positive ones.
+\begin{equation*}
+	\left[\operatorname{Thresh}(\eigvalmat,\nDimensions) \right]_{i,i} = \left\{ \begin{array}{rl}
+	\lambda_{i,i}  & \text{if} \quad i \leq N \; \text{and} \; \lambda_{i,i}>0 \\
+	0  & \text{if} \quad i > N \; \text{or} \; \lambda_{i,i}\leq 0
+	\end{array}\right.
+\end{equation*}
+where the eigenvalues in $\eigvalmat$ are ordered from the largest to the smallest. Following this step, at most $\nDimensions$ eigenvalues of the Gram matrix are different from zero,
+\item factorizing the projected Gram matrix into the product $\Gram=\Transpose{\Dic}\Dic$ by setting:
+\begin{equation}
+\Dic = \eigvalmat^{1/2}\Transpose{\eigvecmat}.
+\end{equation}
+\end{enumerate}
+\item \emph{Dictionary rotation}. Rotate the dictionary $\Dic$ to align it to the training set by solving the problem:
+\begin{equation}\label{eq:rot}
+	\optimal{\orthmat} = \Minimise{\orthmat \Transpose{\orthmat} = \Matrix{I}}{\norm{\Feas - \orthmat\Dic\Coeff}{\Label{F}}}.
+\end{equation}
+The optimal rotation matrix can be calculated by:
+\begin{enumerate}
+\item computing the sample covariance between the observed signals and their approximations $\CovMat \definition (\Dic\Coeff)\Transpose{\Feas}$,
+\item calculating the \Acronym{svd} of the covariance $\CovMat = \Matrix{U}\Matrix{\Sigma}\Transpose{\Matrix{V}}$,
+\item setting the optimal rotation matrix to $\optimal{\orthmat}=\Matrix{V}\Transpose{\Matrix{U}}$,
+\item rotating the dictionary $\Dic \leftarrow \optimal{\orthmat}\Dic$.
+\end{enumerate}
+\end{itemize}
+
+More details about the \Acronym{ipr} algorithm can be found in  \cite{Barchiesi2013Le}, including details of its computational cost.
+%The code of the \Acronym{ipr} method is illustrated in Algorithm \ref{algo:ipr}.
+%
+%\begin{algo}
+%\KwIn{$\Feas, \Dictionary, \Coeff, \coherence_{0}, \nIter$}
+%\KwOut{$\optimal{\Dictionary}$}
+%$\iter\gets1$\;
+%\While{$\iter \leq \nIter$ and $\coherence(\Dictionary)>\coherence_{0}$}{
+%	\tcp{Calculate Gram matrix}
+%	$\Gram \gets \Transpose{\Dictionary}\Dictionary$\;
+%	\tcp{Project ont structural c.s.}
+%	$\operatorname{diag(\Gram)} \gets \Vector{1}$\;
+%	$\Gram \gets \operatorname{Limit}(\Gram,\coherence_{0})$\;
+%	\tcp{Project Gram matrix onto spectral c.s. and factorize}
+%	$[\eigvecmat, \eigvalmat] \gets \Acronym{evd}(\Gram)$\;
+%	$\eigvalmat \gets \operatorname{Thresh}(\eigvalmat,\nDimensions)$\;
+%	$\Dictionary \gets \eigvalmat^{1/2}\Transpose{\eigvecmat}$\;
+%	\tcp{Rotate dictionary}
+%	$\CovMat \gets \Feas\Transpose{(\Dictionary\Coeff)}$\;
+%	$[\Matrix{U},\Matrix{\Sigma},\Matrix{V}] \gets \Acronym{svd}(\CovMat)$\label{algo:ipr:svd}\;
+%	$\OrthMat \gets \Matrix{V}\Transpose{\Matrix{U}}$\;
+%	$\Dictionary \gets \OrthMat\Dictionary$\;
+%	$\iter\gets\iter+1$\;
+%}
+%\caption{\label{algo:ipr}Iterative projections and rotations (\Acronym{ipr})}
+%\end{algo}
+
+\section{Learning Incoherent Subspaces}\label{sec:lis}
+The \Acronym{ipr} algorithm learns a dictionary where all the atoms are mutually incoherent.
+Therefore, given any two disjoint sets $\Lambda\bigcap\Gamma=\emptyset$ that identify non-overlapping collections of atoms, the sub-dictionaries $\Dic_{\Lambda}, \Dic_{\Gamma}$ are also mutually incoherent.
+
+Starting from this observation, the main intuition driving the development of a supervised \Acronym{ipr} (\Acronym{s-ipr}) algorithm for classification is to learn mutually incoherent sub-dictionaries that approximate features from different classes of signals.
+The sub-dictionaries are in turn used to define incoherent sub-spaces, and features are projected onto these sub-spaces yielding discriminative dimensionality reduction.
+
+\subsection{The supervised \Acronym{ipr} algorithm}\label{sec:iprclass}
+Let $\curlyb{\cat_{\iFea}\in\Cat}_{\iFea=1}^{\nFea},\; \Cat=\curlyb{\uniCat_{1},\uniCat_{2},\dots,\uniCat_{\nCat}}$ be a set of labels that identify the category of the vectors of features $\fea_{\iFea}$, whose elements belong to a set $\Cat$ of $\nCat$ possible categories.
+The columns of the matrix $\Feas_{\iCat}$ contain a selection of the features extracted from signals belonging to the $\iCat$-th category.
+
+To learn incoherent sub-dictionaries from the entire set of features, we must first cluster the atoms to different classes\footnote{Note that the term \emph{cluster} implies that a this stage the algorithm needs to make an unsupervised decision, since there is no any a-priori reason to assign a given atom to any particular class.}, and then only proceed with their de-correlation  if they are assigned to different categories (while allowing coherent atoms to approximate features from the same class).
+To this aim, we employ the matrix $\Coeff$ to measure the contribution of every atom to the approximation of features belonging to each class.
+
+Let $\coeff_{\iCat}^{\iAto}$ indicate the $\iAto$-th row of the matrix $\Coeff_{\iCat}$ containing the coefficients that contribute to the approximation of $\Feas_{\iCat}$, and $\nDim_{\iCat}$ indicate the number of its elements.
+A coefficient $\gamma_{\iAto,\iCat}$ is defined as:
+\begin{equation}
+	\gamma_{\iAto,\iCat} \definition \frac{1}{\nDim_{\iCat}}\norm{\coeff_{\iCat}^{\iAto}}{1},
+\end{equation}
+and every atom $\atom_{\iAtom}$ is associated with the category to which it maximally contributes $\optimal{\iCat}_{\iAto} = \underset{\iCat}{\arg\max}\curlyb{\gamma_{\iAto,\iCat}}$.
+
+Grouping together atoms that have been assigned to the same class leads to a set of sub-dictionaries whose size and rank depends on the number of atoms for each class, and to their linear dependence.
+As a general heuristic, if  features corresponding to different classes do not occupy the same sub-space (according to the active elements in $\Coeff$), a full-rank dictionary $\Dic$ with $\nAto \geq \nDim \gg \nCat$ ensures that $\optimal{\iCat}_{\iAto}$ identify $\nCat$ non-empty and disjoint sub-dictionaries $\curlyb{\Dic_{\iCat}}_{\iCat=1}^{\nCat}$.
+
+Once the atoms have been clustered, the Gram matrix $\Gram$ is computed and iteratively projected as in the method described in Section \ref{sec:ipr}, with the difference that equation \eqref{eq:pscs} is modified in order to only constraint the mutual coherence between atoms assigned to different categories
+\small
+\begin{equation}\label{eq:pscs2}
+\operatorname{Limit}({\gram}_{i,j},\coherence_{0},\optimal{\Vector{\iCat}}) = \left\{ \begin{array}{rl}
+	\gram_{i,j}  & \text{if} \quad |\gram_{i,j}|\leq \coherence_{0} \, \text{or} \, \optimal{\iCat}_{i}=\optimal{\iCat}_{j} \\
+	\operatorname{sgn}(\gram_{i,j}) \coherence_{0}  & \text{if} \quad |\gram_{i,j}|   >  \coherence_{0} \, \text{and} \, \optimal{\iCat}_{i}\neq\optimal{\iCat}_{j}
+\end{array} \right.
+\end{equation}
+\normalsize
+
+A further modification of the standard \Acronym{ipr} algorithm presented in \cite{Barchiesi2013Le} consists in the update of the Gram matrix, performed by computing its element-wise average with the projection $\stcmat = \Projection_{\stcset}(\Gram)$ (rather than by using the projection alone).
+This heuristic has led to improved empirical results by preventing $\Gram$ from changing too abruptly.
+
+The complete supervised \Acronym{s-ipr} method is summarised in Algorithm \ref{algo:sipr}.
+Note that the mutual coherence $\coherence_{\optimal{\iCat}}(\Dic) = \underset{\optimal{\iCat}_{i}\neq\optimal{\iCat}_{j}}{\arg\max}\inner{\atom_{i}}{\atom_{j}}$ indicated in this algorithm measures the inner product between any two atoms assigned to different categories since atoms assigned to the same category are allowed to be mutually coherent.
+\begin{algo}
+\KwIn{$\Feas, \Dictionary, \Coeff, \coherence_{0}, \cats, \nIter$}
+\KwOut{$\optimal{\Dictionary}$}
+$\iter\gets1$\;
+\tcp{Cluster atoms}
+$\Coeff_{\iCat} \gets \squareb{\coeff_{j}} \forall j \in \uniCat_{\iCat}$\;
+$\gamma_{\iAto,\iCat} \gets \norm{\coeff_{\iCat}^{\iAto}}{1}/\nDim_{\iCat}$\;
+$\optimal{\iCat}_{\iAto} = \underset{\iCat}{\arg\max}\curlyb{\gamma_{\iAto,\iCat}}$\;
+\While{$\iter \leq \nIter$ and $\coherence_{\optimal{\iCat}}(\Dictionary)>\coherence_{0}$}{
+	\tcp{Calculate Gram matrix}
+	$\Gram \gets \Transpose{\Dictionary}\Dictionary$\;
+	\tcp{Project onto structural c.s.}
+	$\operatorname{diag}(\stcmat) \gets \Vector{1}$\;
+	$\stcmat \gets \operatorname{Limit}(\Gram,\coherence_{0},\optimal{\Vector{\iCat}})$\;
+	$\Gram \gets \frac{1}{2}\Gram + \frac{1}{2}\stcmat$\;
+	\tcp{Project onto spectral c.s. and factorize}
+	$[\eigvecmat, \eigvalmat] \gets \Acronym{evd}(\Gram)$\;
+	$\eigvalmat \gets \operatorname{Thresh}(\eigvalmat,\nDimensions)$\;
+	$\Dictionary \gets \eigvalmat^{1/2}\Transpose{\eigvecmat}$\;
+	\tcp{Rotate dictionary}
+	$\CovMat \gets \Feas\Transpose{(\Dictionary\Coeff)}$\;
+	$[\Matrix{U},\Matrix{\Sigma},\Matrix{V}] \gets \Acronym{svd}(\CovMat)$\label{algo:ipr:svd}\;
+	$\OrthMat \gets \Matrix{V}\Transpose{\Matrix{U}}$\;
+	$\Dictionary \gets \OrthMat\Dictionary$\;
+	$\iter\gets\iter+1$\;
+}
+\caption{\label{algo:sipr}Supervised \Acronym{ipr}}
+\end{algo}
+\subsection{Classification via incoherent subspaces}\label{sec:class}
+The \Acronym{s-ipr} algorithm allows to learn a set of sub-dictionaries $\curlyb{\Dic_{\iCat}}$ that contain mutually incoherent atoms.
+These cannot be directly used to define discriminative subspaces because, depending on $\nDim$ and on the rank of each sub-dictionary, atoms belonging to disjoint sub-dictionaries might span identical subspaces.
+Instead, we fix a rank $\nDimSub\leq\floor{\nDim/\nCat}$ and choose a collection of $\nDimSub$ linearly independent atoms from each sub-dictionary $\Dic_{\iCat}$, using the largest values of $\gamma_{\iAto,\iCat}$ to define a picking order.
+Thus, we obtain a set $\curlyb{\Spa_{\iCat}}_{\iCat=1}^{\nCat}$ of incoherent sub-spaces of rank $\nDimSub$ embedded in the space $\ambient^{\nDim}$, and use them to derive a feature transform for classification.
+
+Each feature vector $\fea_{\iFea}$ that belongs to the class $\cat_{\iFea}$ is projected onto the relative subspace, yielding a set of transformed features $\curlyb{\newFea_{\iFea}}_{\iFea=1}^{\nFea}$.
+\begin{equation}
+\newFea_{\iFea} = \Spa_{\cat_{\iFea}}\pseudoinverse{\Spa_{\cat_{\iFea}}}\fea_{\iFea}
+\end{equation}
+where $\Spa^{\dagger}$ denotes the Moore-Penrose pseudo-inverse of the matrix $\Spa$ and needs to be used in place of the transposition operator because the columns of $\Spa$ are in general not orthogonal.
+
+When an unlabelled signal is presented to the classifier, the corresponding vector of features $\fea$ is projected onto all the learned sub-spaces.
+Then, the nearest sub-space is chosen using an Euclidean distance measure, and the corresponding projection $\newFea$ used as the transformed feature.
+\begin{align}
+	\optimal{\iCat} &= \underset{\iCat}{\arg\min}\norm{\fea-\Spa_{\iCat}\pseudoinverse{\Spa_{\iCat}}\fea}{2}	\\
+	\newFea &= \Spa_{\optimal{\iCat}}\pseudoinverse{\Spa_{\optimal{\iCat}}}\fea
+\end{align}
+The subspace $\optimal{\iCat}$ can be directly used as an estimator of the category of the signal $\optimal{\cat}$.
+Alternatively, a simple \emph{k-neaerst neighbour} classifier can be employed on the transformed features, and a class can be inferred as:
+\begin{equation}
+\optimal{\cat} = \texttt{knn}(\newFea,\NewFeas,\cats)
+\end{equation}
+where $\NewFeas$ represents the matrix of training features after the transform stage.
+This latter approach is especially suitable when working with a large number of classes in a space of relatively small dimension (as in the numerical experiment presented in Section \ref{sec:fisher} where $\nCat=3$ and $\nDim=3$), as in this case multiple classes might be assigned to the same subspace.
+\section{Numerical Experiments}\label{sec:ne}
+\subsection{Classifying the \Acronym{fisheriris} dataset}\label{sec:fisher}
+To illustrate the \Acronym{s-ipr} algorithm for feature transform, we apply it to the $\Acronym{fisheriris}$ dataset \cite{Fisher1936Th}.
+This contains data for the classification of $150$ iris specimens into the classes \emph{setosa}, \emph{versicolor} and \emph{virginica}.
+The features corresponding to measurements on the sepal length, sepal width and petal length for the flowers are stored in the matrix $\Feas\in\ambient^{3\times 150}$.
+As a pre-processing step, we subtract the mean and normalise the standard deviation of the features to avoid large variations or offsets in any of the dimensions of the data.
+
+The dictionary learning is run using the \Acronym{SMALLBox} toolbox and the Incoherent dictionary learning add-on\footnote{http://code.soundsoftware.ac.uk/projects/smallbox.}.
+We learn an over-complete dictionary $\Dic$ consisting of $\nAto=10$ atoms and a sparse approximation $\Coeff$ that uses $\nActiveAtoms=2$ active atoms for each signal.
+Incoherent sub-dictionaries are learned by the \Acronym{s-ipr} method, setting the mutual coherence to the theoretical lower bound attainable by a $\nDim\times \nAto$ dictionary $\mu_{\Label{min}} = \sqrt{(\nAto-\nDim)/\nDim(\nAto-1)}\approx 0.5$ \cite{Strohmer2003Gr}.
+A maximum number of $\nIter=20$ iterations is set as the stopping criterion.
+These parameters are within the range used in previous studies of incoherent dictionary learning \cite{Barchiesi2013Le}, and have not been optimised for this task.
+
+The proposed method is used to learn $\nDimSub=1$ dimensional sub-spaces, and is  compared to \Acronym{pca} with a number of components $\nComponents=2$ and to \Acronym{lda}.
+Figure \ref{fig:fisher} shows the results of the experiment.
+The original features show a clear cluster corresponding to the setosa class, and overlapping data for the other two classes.
+Applying \Acronym{pca} reduces the dimensionality of the $3$-dimensional features to a rank $2$ principal sub-space, but worsens the separation between classes.
+Both \Acronym{lda} and the proposed method are able to improve the separation of the three classes, but in remarkably different ways: while the former optimises the linear separability of different classes, features derived from \Acronym{s-ipr} are projected onto low-dimensional discriminative subspaces.
+
+To assess the classification performance of the various techniques, we created a random $5$-fold partition of the training set used for cross-validation, and classified the transformed features using \Acronym{knn} with $5$ neighbours.
+The misclassification rate $\Acronym{mcr}$ defined as the number of misclassified samples divided by the total number of samples is displayed for the various methods in the first row of Table \ref{tab:fisher}.
+\Acronym{s-ipr} and \Acronym{lda} are both able to attain a $7\%$ misclassification error, improving on the $10\%$ obtained on the original features.
+The worst result is achieved by \Acronym{pca}, whose dimensionality reduction has the effect of mixing together features belonging to different classes, and achieves an error of $44\%$.
+\begin{figure*}[!tb]
+\begin{center}
+%\includegraphics[width=\textwidth]{fisher3d.pdf}
+\caption{\label{fig:fisher}Features transform for the fisheriris dataset. Red circular markers correspond to setosa, green square markers to versicolor and blue diamond markers to virginica.}
+\end{center}
+\end{figure*}
+\begin{table}[!htb]
+\begin{center}
+\begin{tabular}{|l|c|c|c|c|}
+\hline
+  & none & \Acronym{pca} & \Acronym{lda} & \Acronym{s-ipr} \\
+  \hline
+  \textbf{\Acronym{mcr} - \Acronym{fisheriris}} & 0.10 & 0.44 & 0.07 & 0.07\\
+\textbf{\Acronym{mcr} - \Acronym{usps}} & 0.343 & 0.433 & 0.359 & 0.315 \\
+\hline
+\end{tabular}
+\caption{\label{tab:fisher} Misclassification error evaluated using different feature transform methods on the \Acronym{fisheriris} and \Acronym{usps} datasets (none indicates no feature transform).}
+\end{center}
+\end{table}
+\subsection{Classification of the \Acronym{usps} digits dataset}
+In order to evaluate the proposed method on a more challenging dataset that  contains signals of higher dimension, we run a similar experiment on the \Acronym{usps} digits dataset, which consists of a collection of $16\times 16$ pixels images of hand-written digits \footnote{http://www.cs.nyu.edu/~roweis/data.html.}.
+We selected the digits \Acronym{1}, \Acronym{3} and \Acronym{8} resulting in a total of $1405$ examples stored in the matrix $\Feas\in\ambient^{256\times 1405}$.
+
+We investigated the effect of the parameters \\ $\nAto=\curlyb{256,512,1024}$, $\nDimSub=\curlyb{2,4,8,16,32,64,128}$ and $\nActiveAtoms = \curlyb{2,25,128,256}$ on the misclassification error, and compared the \Acronym{s-ipr} to \Acronym{pca} and \Acronym{lda}.
+For \Acronym{pca}, the number of principal components $\nComponents$ is automatically set by the algorithm to choose an approximation that retains $95\%$ of the variance of the data.
+The second row of Table \ref{tab:fisher} shows the results obtained with the best choice of parameters $\optimal{\nAto}=512$ (corresponding to a 2 times over-complete dictionary), $\optimal{\nDimSub}=64$ and $\optimal{\nActiveAtoms}=128$.
+In this case, \Acronym{s-ipr} achieves a $31.5\%$ misclassification error, while the other two techniques do not improve the \Acronym{k-nn} classification on to the original features, which scores $34.3\%$.
+
+Analysing the \Acronym{mcr} as a function of the parameters $\nAto$, $\nDimSub$ and $\nActiveAtoms$, the most relevant trend observed empirically on this dataset is that a value $\nDimSub<16$ steeply increases the misclassification error.
+This suggests that the dimensionality of the incoherent sub-spaces must be large enough to retain discriminative characteristics of the original high-dimensional data.
+The dependency of \Acronym{mcr} on the other two parameters is less significant, suggesting that dictionary learning leads to discriminative subspaces regardless of the number of active atoms $\nActiveAtoms$ or the number of atoms in the dictionary $\nAto$.
+\section{Conclusion}\label{sec:end}
+We have presented the \Acronym{s-ipr} algorithm for learning incoherent subspaces, and employed it as a feature transform method in the context of supervised classification.
+The encouraging experimental results obtained on the \Acronym{fisheriris} and the \Acronym{usps} datasets suggest that the proposed algorithm can overcome limitations of standard feature transforms methods, making it a viable tool for machine learning research and practice.
+A better theoretical understanding of the performance of \Acronym{s-ipr} (especially in relation to other methods for subspace and manifold learning), and further validations of the technique on challenging classification problems constitute the most promising avenues for future research.
+
+% References should be produced using the bibtex program from suitable
+% BiBTeX files (here: strings, refs, manuals). The IEEEbib.bst bibliography
+% style file from IEEE produces unsorted bibliography list.
+% -------------------------------------------------------------------------
+\bibliographystyle{IEEEbib}
+\bibliography{bibliography.bib}
+
 \end{document}

### 2014-01-13 manuscript.tex
added figures 1, 2 and 3

@@
 \subsubsection{\Acronym{pca}}
+
 Let $\curlyb{\fea_{\iFea} \in \real^{\nDim}}_{\iFea=1}^{\nFea}$ be a set of vectors containing features extracted from $\nFea$ training signals.
@@
 The bases are optimised from the data to identify their principal components, that is, the sub-spaces that retain the maximum variance of the features.
-First, the eigenvalue decomposition of the outer product $\Matrix{X}\Transpose{\Matrix{X}} = \Matrix{Q}\Matrix{\Lambda}\Transpose{\Matrix{Q}}$ is computed.
-Then, the largest $\nComponents$ eigenvalues are selected from the matrix $\Matrix{\Lambda}$, and the corresponding eigenvectors are
+
+To compute the dictionary, the eigenvalue decomposition of the outer product
+\begin{equation}
+\Matrix{X}\Transpose{\Matrix{X}} = \Matrix{Q}\Matrix{\Lambda}\Transpose{\Matrix{Q}}
+\end{equation}
+is first calculated.
+Then, the $\nComponents$ eigenvectors corresponding to the $\nComponents$ largest eigenvalues are selected from the matrix $\Matrix{Q}$, and scaled to unit $\ell_{2}$ norm to form the dictionary $\Dic$.
 
@@
 In contrast, \Acronym{lda} is a supervised method for feature transform whose objective is to explicitly maximise the separability of classes in the transformed domain.
-The within-classes scatter matrix is defined to measure how features belonging to the same class are clustered around their mean vector (the vector obtained by averaging all observations associated with a given category), and the  between-classes scatter matrix is defined to measure the distances between the mean vectors.
-A square linear transform $\Matrix{M} \in \real^{\nDim\times\nDim}$ is optimised to obtain a new set of features $\newFea_{\Acronym{lda}}=\Matrix{M}\fea$ according to an objective that promotes features belonging to the same class to be near each other (according to a Euclidean distance), and far apart from features belonging to different classes.
-\Acronym{lda} explicitly seeks to enhance the discriminative power of features, but does not perform dimensionality reduction.
 
+Let $\obsSet_{\iUniCat}$ be a set indexing features extracted from data belonging to the $\iUniCat$-th category, let
+\begin{equation}
+\average{\fea}_{\iUniCat} \definition \frac{1}{\abs{\obsSet_{\iUniCat}}}\sum_{\iObs\in\obsSet_{\iUniCat}} \fea_{\iObs}
+\end{equation} 
+be the $\iUniCat$-th class feature centroid, and $\average{\fea}\definition\sum_{\iObs=1}^{\nObs}\fea_{\iObs}$ the centroid of the features extracted from the entire training dataset.
+The between-classes scatter matrix
+\begin{equation}
+\Matrix{S_{b}} \definition \sum_{\iUniCat=1}^{\nUniCat}\abs{\obsSet_{\iUniCat}}\roundb{\average{\fea}_{\iUniCat}-\average{\fea}}\Transpose{\roundb{\average{\fea}_{\iUniCat}-\average{\fea}}}
+\end{equation} 
+is defined to measure the mutual distances between the centroids of different classes, while the within-classes scatter matrix 
+\begin{equation}
+\Matrix{S}_{w}\definition \sum_{\iUniCat=1}^{\nUniCat}\sum_{\iObs\in\obsSet_{\iUniCat}}\roundb{\fea_{\iObs}-\average{\fea}_{\iUniCat}}\Transpose{\roundb{\fea_{\iObs}-\average{\fea}_{\iUniCat}}}
+\end{equation}
+quantifies the distances between features belonging to the same class.
+To maximise an objective function $\objective(\Matrix{W})\definition\frac{\abs{\Transpose{\Matrix{W}}\Matrix{S}_{b}\Matrix{W}}}{\abs{\Transpose{\Matrix{W}}\Matrix{S}_{w}\Matrix{W}}}$ that promotes features belonging to the same class to be near each other and far away from features belonging to other classes, the eigenvalue decomposition of the matrix
+\begin{equation}
+\pseudoinverse{\Matrix{S}_{w}}\Matrix{S}_{b} = \Matrix{Q}\Matrix{\Lambda}\Transpose{\Matrix{Q}}
+\end{equation}
+is computed, and the features $\fea$ are projected onto the space spanned by its $(\nUniCat-1)$ eigenvectors corresponding to the largest $(\nUniCat-1)$ eigenvalues.
+
+\Acronym{lda} explicitly seeks to enhance the discriminative power of features by optimising the objective $\objective$.
 \subsection{Supervised dimensionality reduction}
@@
 
-In practice, \Acronym{s-pca} differs from \Acronym{pca} in that it calculates the eigenvalue decomposition of a matrix $\Matrix{Q}$ defined as follows:
+In practice, \Acronym{s-pca} differs from \Acronym{pca} in that it calculates the eigenvalue decomposition of a matrix $\Matrix{R}$ defined as follows:
 \begin{equation}
-\Matrix{Q} \definition \Matrix{X}\Matrix{H}\Matrix{L}\Matrix{H}\Transpose{\Matrix{X}}
+\Matrix{R} \definition \Matrix{X}\Matrix{H}\Matrix{L}\Matrix{H}\Transpose{\Matrix{X}}
 \end{equation}
-were $\Matrix{H} \definition \Matrix{I} - \\Vector{e}\Transpose{\Vector{e}}$ is a \emph{centring} matrix and $\Matrix{L} \definition \Vector{\cat}\Transpose{\Vector{\cat}}$ is the kernel matrix of the class variable that is constructed by computing the outer product of the vectors resulting from assigning different numerical values to each category.
+were $\Matrix{H} \definition \Matrix{I} - \Vector{e}\Transpose{\Vector{e}}$ is a \emph{centring} matrix and $\Matrix{L} \definition \Vector{\cat}\Transpose{\Vector{\cat}}$ is the kernel matrix of the class variable that is constructed by computing the outer product of the vectors resulting from assigning different numerical values to each category.
+
+Other methods for supervised dimensionality reduction include metric learning algorithms \cite{xing2002distance}, sufficient dimensionality reduction \cite{li1991sliced} and supervised principal components \cite{Bair06predictionby}.
 
@@
 \section{Numerical Experiments}\label{sec:ne}
+\subsection{Feature visualisation}
+To illustrate the \Acronym{s-ipr} algorithm for feature transform, we first run visualisation experiments depicting how different feature transform methods act on training and test data.
+\subsubsection{Synthetic data}
+Figure \ref{fig:toy} displays a total of $1500$ synthetic features in $\real^{2}$ belonging to $3$ different classes that we generated for this experiment.
+For each class, first we draw values distributed uniformly in the interval $\squareb{-1,1}$ and assign them to the first component of the features (the \emph{x} coordinate).
+Then, we add Gaussian noise with variance $0.1$ to the second component (the \emph{y} coordinate), and we rotate the resulting data by the angles $\theta_{0}=0$, $\theta_{1}=\pi/4$ and $\theta_{3}=\pi/2$ for the $3$ classes respectively.
+This way, features belonging to different classes are clustered along different one-dimensional sub-spaces of $\real^{2}$.
+\begin{figure}
+\includegraphics[width=\textwidth]{./Code/Datasets/GetToyExampleDataset.pdf}
+\caption{\label{fig:toy}Synthetic data generated along one-dimensional subspaces of $\real^{2}$.}
+\end{figure}
+
+Figure
+\begin{figure}
+\includegraphics[width=\textwidth]{./Code/toyvisu.pdf}
+\caption{\label{fig:toy}Synthetic data generated along one-dimensional subspaces of $\real^{2}$.}
+\end{figure}
+\subsubsection{Iris dataset}
+\begin{figure}
+\includegraphics[width=\textwidth]{./Code/fishervisu.pdf}
+\caption{\label{fig:toy}Synthetic data generated along one-dimensional subspaces of $\real^{2}$.}
+\end{figure}
+
 \subsection{Classifying the \Acronym{fisheriris} dataset}\label{sec:fisher}

### 2014-01-13 code/Datasets/GetToyExampleDataset.m

@@ -1,9 +1,8 @@
-function [fea,cat] = GetToyExampleDataset
+function [fea,cat] = GetToyExampleDataset(visu)
+if ~exist('visu','var'), visu=false; end
+% reset(RandStream.getGlobalStream);    %set random number generator to reproduce results
 
-clear, clc, close all
-reset(RandStream.getGlobalStream);    %set random number generator to reproduce results
-
-nData = 1000;						%number of data for each class
+nData = 500;						%number of data for each class
 theta = [0, pi/4, pi/2];            %angles
 nThetas = length(theta);			%number of angles
 fea = zeros(nThetas*nData,2);
@@ -12,8 +11,12 @@ for iTheta=1:nThetas
     fea((iTheta-1)*nData+1:iTheta*nData,:) = randomsamp(nData,theta(iTheta));
     cat((iTheta-1)*nData+1:iTheta*nData) = iTheta*ones(nData,1);
 end
-% figure
-% gscatter(fea(:,1),fea(:,2),cat,'krb','.',10);
+if visu
+    figure
+    gscatter(fea(:,1),fea(:,2),cat,'krb','.',10);
+    [a,b] = fileparts(which(mfilename));
+    savegraph([a,filesep,b]);
+end
 end
 
 function X = randomsamp(nData,theta)

### 2014-01-13 code/LDAFeaturesTransform.m

@@ -9,8 +9,6 @@ St = Sb + Sw;                               %Total scatter matrix
 J = @(W) trace(pinv(W'*St*W)*(W'*Sb*W));    %as a function of projection matrix
 
 [D,X] = eig(pinv(St)*Sb);                   %compute evd
-idx = diag(real(X))>1e-8;                   %find non-zero eigenvalues
-W = real(D(:, idx));                        %select corresponding eigenvectors
 W = real(D(:,1:length(unique(traCat))-1));
 % assert(size(W,2)==(length(unique(traCat))-1));
 

### 2014-01-13 code/TestTransforms.m

@@ -12,7 +12,7 @@ end
 
 if ~exist('par','var') || isempty(par), par = struct; end
 
-options = statset('UseParallel',true); %set option to use parallel computing
+options = statset('UseParallel',false); %set option to use parallel computing
 
 def.nAto = 2*size(fea,2); %number of atoms (twice overcomplete dictionary)
 def.subSpaRan = 1; %subspace range
@@ -84,26 +84,29 @@ iprMcr = sum(sum(iprConMat-diag(diag(iprConMat))))/sum(sum(iprConMat));
 % J1 = @(fea,cat) trace(pinv(ClassificationDiscriminant.fit(fea,cat).Sigma)*ClassificationDiscriminant.fit(fea,cat).BetweenSigma);
 % J2 = @(fea,cat) trace(ClassificationDiscriminant.fit(fea,cat).BetweenSigma)/trace(ClassificationDiscriminant.fit(fea,cat).Sigma);
 % J3 = @(fea,cat) log(abs(det(ClassificationDiscriminant.fit(fea,cat).BetweenSigma))) - ClassificationDiscriminant.fit(fea,cat).LogDetSigma;
-mets = {'none','IPR'};
+mets = {'none','S-PCA','LDA','S-IPR'};
 if par.visu
     nMet = length(mets);
-    sqrNMet = ceil(sqrt(nMet));
-    % Display features
+    sqrNMet = ceil(sqrt(nMet-1));
+
     close all
     for iMet=1:nMet
         switch mets{iMet}
             case 'none'
+                figure(2)
                 traFea = traFea;
                 mat = conMat;
                 err = mcr;
             case 'PCA'
+                figure(2)
                 if size(pcaTraFea,2)~=3
                     traFea = [pcaTraFea zeros(size(pcaTraFea,3-size(pcaTraFea,2)),1)];
                     tesFea = [pcaTesFea zeros(size(pcaTesFea,3-size(pcaTraFea,2)),1)];
                 end
                 mat = pcaConMat;
                 err = pcaMcr;
-            case 'SPCA'
+            case 'S-PCA'
+                figure(2)
                 if size(spcaTraFea,2)~=3
                     traFea = [spcaTraFea zeros(size(spcaTraFea,1),3-size(spcaTraFea,2))];
                     tesFea = [spcaTesFea zeros(size(spcaTesFea,1),3-size(spcaTraFea,2))];
@@ -111,17 +114,19 @@ if par.visu
[…]

### 2014-01-13 manuscript.tex
classification figures

@@
 \usepackage[boxruled,linesnumbered]{algorithm2e}
-\usepackage{subfig,graphicx}
+\usepackage{subfigure,graphicx}
 \usepackage{setspace}
@@
 \includegraphics[width=\textwidth]{./Code/toyvisu.pdf}
-\caption{\label{fig:toy}Synthetic data generated along one-dimensional subspaces of $\real^{2}$.}
+\caption{\label{fig:toyNewFea}Feature transform applied to the synthetic data in Figure \ref{fig:toy}.
+Different colours correspond to different classes, `+' and `o' markers represent samples taken from the training and test set respectively.}
 \end{figure}
@@
 \begin{figure}
+\includegraphics[width=\textwidth]{./Code/Datasets/GetFisherIrisDataset.pdf}
+\caption{\label{fig:fisher3} First three features of the `iris' dataset depicting measurements of sepal length, sepal with and petal length of three iris species.}
+\end{figure}
+\begin{figure}
 \includegraphics[width=\textwidth]{./Code/fishervisu.pdf}
-\caption{\label{fig:toy}Synthetic data generated along one-dimensional subspaces of $\real^{2}$.}
+\caption{\label{fig:toyNewFea}Feature transform applied to the iris data in Figure \ref{fig:fisher3}.
+Different colours correspond to different classes, `+' and `o' markers represent samples taken from the training and test set respectively.}
+\end{figure}
+
+\subsection{Classification results}
+\begin{figure}
+\centering
+\subfigure{\includegraphics[width=.45\textwidth]{Code/Util/fisheriris.pdf}}
+\subfigure{\includegraphics[width=.45\textwidth]{Code/Util/balance.pdf}}
+\subfigure{\includegraphics[width=.45\textwidth]{Code/Util/parkinsons.pdf}}
+\subfigure{\includegraphics[width=.45\textwidth]{Code/Util/sonar.pdf}}
+\caption{}
 \end{figure}
@@
 The worst result is achieved by \Acronym{pca}, whose dimensionality reduction has the effect of mixing together features belonging to different classes, and achieves an error of $44\%$.
-\begin{figure*}[!tb]
-\begin{center}
-%\includegraphics[width=\textwidth]{fisher3d.pdf}
-\caption{\label{fig:fisher}Features transform for the fisheriris dataset. Red circular markers correspond to setosa, green square markers to versicolor and blue diamond markers to virginica.}
-\end{center}
-\end{figure*}
+
 \begin{table}[!htb]
@@
 
-% References should be produced using the bibtex program from suitable
-% BiBTeX files (here: strings, refs, manuals). The IEEEbib.bst bibliography
-% style file from IEEE produces unsorted bibliography list.
-% -------------------------------------------------------------------------
 \bibliographystyle{IEEEbib}
 \bibliography{bibliography.bib}
-
 \end{document}

### 2014-01-13 code/Datasets/GetFisherIrisDataset.m

@@ -1,9 +1,28 @@
-function [fea,cat] = GetFisherIrisDataset()
+function [fea,cat] = GetFisherIrisDataset(visu,dim)
 % Gets the fisher iris dataset from the web and returns the following variables
 % - fea: matrix containing sepal length, width, petal length and width along the columns
 % - cat: cell containing the classes
+if ~exist('visu','var'), visu = false; end
 url = 'http://archive.ics.uci.edu/ml/machine-learning-databases/iris/iris.data';
 s = urlread(url);
 C = textscan(s,'%f,%f,%f,%f,%s');
 fea = [C{1} C{2} C{3} C{4}];
 cat = C{5};
+
+if visu
+    if ~exist('dim','var'), dim = 3; end
+    fea = fea(:,1:dim);
+    figure
+    if dim==2
+        gscatter(fea(:,1),fea(:,2),cat);
+    elseif dim==3
+        gscatter3(fea(:,1),fea(:,2),fea(:,3),cat);
+    end
+    xlabel('sepal length')
+    ylabel('sepal width')
+    zlabel('petal length')
+    legend('setosa','versicolor','virginica')
+    [a,b] = fileparts(which(mfilename));
+    savegraph([a,filesep,b]);
+end
+

### 2014-01-13 code/TestClassSubSpaRan.m

@@ -12,7 +12,7 @@ if length(strfind([A(:).name],'ompmex'))<3 %if the number of files containing 'o
 end
 
 % parse dataset functions
-datfun = {@GetToyExampleDataset,@GetFisherIrisDataset,@GetBalanceDataset,...
+datfun = {@GetFisherIrisDataset,@GetBalanceDataset,...
     @GetParkinsonsDataset,@GetSonarDataset};
 datasets    = cell(length(datfun),1);
 for i=1:length(datfun)
@@ -37,17 +37,8 @@ for iDat=1:length(datasets)
         fprintf('done\n');
     end
     mcrs = [res(iDat,:).mcr]';
-    figure, PlotMCRs(mcrs(:,2:end),subSpaRanks,datasets{iDat});
+    figure, PlotMCRs(mcrs,subSpaRanks,datasets{iDat});
 end
-
-function PlotMCRs(mcrs,subSpaRanks,dataset)
-methodNames = {'PCA','SPCA','LDA','IPR'};
-plot(subSpaRanks,mcrs,'o--','MarkerSize',10)
-legend(methodNames);
-xlabel('Sub-space rank');
-ylabel('misclassification ratio');
-title(sprintf('Classification of %s dataset',dataset));
-set(gca,'YGrid','on','XGrid','off','XTick',subSpaRanks);
 
 function v = TrimEnd(v)
 % trim the end of a linspace vector if it contains equal values

### 2014-01-13 code/TestTransforms.m

@@ -4,10 +4,10 @@ addpath(genpath(pwd)); %add path to matlab search path
 
 %% Parameters and defaults
 if ~exist('fea','var') || isempty(fea) || ~exist('cat','var') || isempty(cat)
-    [fea,cat] = GetFisherIrisDataset;
+    [fea,cat] = GetToyExampleDataset;
     if isnumeric(cat), cat = cellstr(num2str(cat)); end
-    nDim = 3; %data has 4 dimensions
-    fea = fea(:,1:nDim);
+%     nDim = 3; %data has 4 dimensions
+%     fea = fea(:,1:nDim);
 end
 
 if ~exist('par','var') || isempty(par), par = struct; end
@@ -84,7 +84,7 @@ iprMcr = sum(sum(iprConMat-diag(diag(iprConMat))))/sum(sum(iprConMat));
 % J1 = @(fea,cat) trace(pinv(ClassificationDiscriminant.fit(fea,cat).Sigma)*ClassificationDiscriminant.fit(fea,cat).BetweenSigma);
 % J2 = @(fea,cat) trace(ClassificationDiscriminant.fit(fea,cat).BetweenSigma)/trace(ClassificationDiscriminant.fit(fea,cat).Sigma);
 % J3 = @(fea,cat) log(abs(det(ClassificationDiscriminant.fit(fea,cat).BetweenSigma))) - ClassificationDiscriminant.fit(fea,cat).LogDetSigma;
-mets = {'none','S-PCA','LDA','S-IPR'};
+mets = {'PCA','S-PCA','LDA','S-IPR'};
 if par.visu
     nMet = length(mets);
     sqrNMet = ceil(sqrt(nMet-1));
@@ -102,6 +102,9 @@ if par.visu
                 if size(pcaTraFea,2)~=3
                     traFea = [pcaTraFea zeros(size(pcaTraFea,3-size(pcaTraFea,2)),1)];
                     tesFea = [pcaTesFea zeros(size(pcaTesFea,3-size(pcaTraFea,2)),1)];
+                else
+                    traFea = pcaTraFea;
+                    tesFea = pcaTesFea;
                 end
                 mat = pcaConMat;
                 err = pcaMcr;
@@ -151,7 +154,7 @@ mar = {'o','o','o','x','^','>'};
 size = 10;
 for iCat=1:length(uniCat) %for every category
     ind = strcmp(uniCat(iCat),traCat); %find indexes of data in the trainig set belonging to iCat category
[…]

### 2014-01-13 code/TestClassSubSpaRan.m
minor edits

@@ -29,7 +29,7 @@ for iDat=1:length(datasets)
     if ~isfloat(fea), fea = double(fea); end    %transform features to float if needed
     fprintf('done\n');
     feaDim = size(fea,2);
-    subSpaRanks = TrimEnd(ceil(linspace(1,feaDim,5)));
+    subSpaRanks = rmd(ceil(linspace(1,feaDim,20)));
     for iSSR = 1:length(subSpaRanks)
         fprintf('testing transforms using subspaces of rank %d... ',subSpaRanks(iSSR));
         par.subSpaRan = subSpaRanks(iSSR);
@@ -38,12 +38,5 @@ for iDat=1:length(datasets)
     end
     mcrs = [res(iDat,:).mcr]';
     figure, PlotMCRs(mcrs,subSpaRanks,datasets{iDat});
-end
-
-function v = TrimEnd(v)
-% trim the end of a linspace vector if it contains equal values
-ind = find(v==max(v));
-if length(ind) > 1
-    v(ind(2:end)) = [];
 end
 

### 2014-01-13 code/Util/PlotMCRs.m
plot functions

@@ -1,0 +1,15 @@
+function PlotMCRs(mcrs,subSpaRanks,dataset)
+methodNames = {'none','LDA','PCA','S-PCA','S-IPR'};
+plot(subSpaRanks(end),mean(mcrs(:,1)),'ks','MarkerSize',10)
+hold on
+plot(subSpaRanks(end)-1,mean(mcrs(:,4)),'rs','MarkerSize',10)
+plot(subSpaRanks,mcrs(:,[1,2,5]),'--o','MarkerSize',10)
+legend(methodNames);
+xlabel('Sub-space rank');
+ylabel('misclassification ratio');
+title(sprintf('Classification of %s dataset',dataset));
+set(gca,'YGrid','on','XGrid','off',...
+    'XTick',rmd([subSpaRanks(1:end-1), subSpaRanks(end)-1, subSpaRanks(end)]));
+a = fileparts(which(mfilename));
+savegraph([a filesep dataset])
+saveas(gcf,[dataset '.fig'])

### 2014-01-13 code/Util/gscatter3.m

@@ -1,0 +1,9 @@
+function hh = gscatter3(x,y,z,cat)
+% 3d group scatter
+uniCat = unique(cat);
+nUniCat = length(uniCat);
+for iUniCat = 1:nUniCat
+    ind = find(strcmp(cat,uniCat(iUniCat)));
+    hh = scatter3(x(ind,:),y(ind,:),z(ind,:));
+    hold on
+end

### 2014-01-13 code/Util/rmd.m

@@ -1,0 +1,9 @@
+function v = rmd(w)
+v(1) = w(1);
+c = 2;
+for i=2:length(w)
+    if w(i) > w(i-1)
+        v(c) = w(i);
+        c = c + 1;
+    end
+end

### 2014-01-15 manuscript.tex
added results figures

@@
 %%%%%%%%%%%%%%%%%%%%%%PACKAGES%%%%%%%%%%%%%%%%%%%%%
-\usepackage{amsmath,graphicx}
+\usepackage{amsmath}
 \usepackage[english]{babel}
@@
 \usepackage[T1]{fontenc}
-\usepackage{epsfig}
 \usepackage{amsfonts,amssymb,amsmath,amsthm,bm}
@@
 \usepackage{setspace}
+\usepackage{hyperref}
 \usepackage{enumitem}
+\usepackage{graphicx}
+\usepackage{epstopdf}
 
@@
 \begin{abstract}
-
+In this article we present the supervised iterative projections and rotations (\Acronym{s-ipr}) algorithm, a method for learning discriminative incoherent subspaces from data.
+We derive \Acronym{s-ipr} as a supervised extension of our previously developed iterative projections and rotations (\Acronym{ipr}) algorithm for incoherent dictionary learning, and we employ it to learn incoherent sub-spaces that model signals belonging to different classes.
+We test our method as a feature transform for supervised classification, first by visualising transformed features from a synthetic dataset and from the `iris' dataset, then by using the resulting features in a classification experiment.
+While the visualisation results are promising, we find that \Acronym{s-ipr} generally performs worse than traditional and state-of-the-art techniques for supervised dimensionality reduction in terms of the misclassification ratio.
 \end{abstract}
@@
 Then, the $\nComponents$ eigenvectors corresponding to the $\nComponents$ largest eigenvalues are selected from the matrix $\Matrix{Q}$, and scaled to unit $\ell_{2}$ norm to form the dictionary $\Dic$.
-
-Let $\Spa\definition\Dic_{1:\nComponents}\in\ambient^{\nDim\times\nComponents}$ denote the sub-dictionary constructed from the $\nComponents<\nDim$ principal components of the dataset.
-A new set of transformed features $\newFea_{\Acronym{pca}} = \Spa\Transpose{\Spa}\fea$ is computed by projecting the data onto the sub-space spanned by the columns of $\Dic_{1:\nComponents}$ (that is, onto the $\nComponents$-dimensional principal sub-space).
+A new set of transformed features $\newFea_{\Acronym{pca}} = \Dic\Transpose{\Dic}\fea$ is computed by projecting the data onto the sub-space spanned by the columns of $\Dic$ (that is, onto the $\nComponents$-dimensional principal sub-space).
 This operation reduces the dimensionality of the features by projecting them onto a linear subspace embedded in $\real^{\nDim}$.
@@
 \Acronym{lda} explicitly seeks to enhance the discriminative power of features by optimising the objective $\objective$.
-\subsection{Supervised dimensionality reduction}
-\subsubsection{Supervised PCA}
+\subsection{Supervised \Acronym{pca}}
 Related works that extend \Acronym{pca} include the supervised \Acronym{pca} (\Acronym{s-pca}) proposed by Barshan et al.
@@
 \Acronym{s-pca} is based on the theory of reproducing kernel Hilbert spaces (\Acronym{rkhs}) (that are spaces of functions which satisfy certain properties and map elements from an arbitrary set to the set of complex numbers) \cite{Aronszajn:1950}, and on the so-called Hilbert-Schmidt independence criterion (\Acronym{hsic})\cite{gretton2005measuring}.
-The \Acronym{hsic} is used to estimates the statistical dependence of two random variables based on the fact that this quantity is related to the correlation of functions belonging to their respective \Acronym{rkhs}.
+The \Acronym{hsic} is used to estimate the statistical dependence of two random variables based on the fact that this quantity is related to the correlation of functions belonging to their respective \Acronym{rkhs}.
 While \Acronym{hsic} is defined in terms of the probability density function of the two random variables,  empirical estimates of \Acronym{hsic} can be obtained from finite sequences of their realisations.
@@
 \end{equation}
-were $\Matrix{H} \definition \Matrix{I} - \Vector{e}\Transpose{\Vector{e}}$ is a \emph{centring} matrix and $\Matrix{L} \definition \Vector{\cat}\Transpose{\Vector{\cat}}$ is the kernel matrix of the class variable that is constructed by computing the outer product of the vectors resulting from assigning different numerical values to each category.
+were $\Matrix{H} \definition \Matrix{I} - \Vector{e}\Transpose{\Vector{e}}$ is a so-called \emph{centring} matrix\footnote{Here $\Vector{e}$ is a vector of ones.} and $\Matrix{L} \definition \Vector{\cat}\Transpose{\Vector{\cat}}$ is the kernel matrix of the class variable that is constructed by computing the outer product of the vectors resulting from assigning different numerical values to each category.
 
-Other methods for supervised dimensionality reduction include metric learning algorithms \cite{xing2002distance}, sufficient dimensionality reduction \cite{li1991sliced} and supervised principal components \cite{Bair06predictionby}.
+\subsection{Other related work}
+The union of incoherent sub-spaces model proposed by Schnass and Vandergheynst \cite{Schnass2010A-} employes a very similar intuition to the one that inspired our proposed method, and models features belonging to different classes using incoherent subspaces.
 
-\subsection{Incoherent subspace learning and other related work}
+Other methods for supervised dimensionality reduction include metric learning algorithms \cite{xing2002distance}, sufficient dimensionality reduction \cite{li1991sliced} and Bair's supervised principal components \cite{Bair06predictionby}.
 
-and methods for manifold learning reviewed by Van Der Maaten et al.
+Manifold learning techniques are used to model nonlinear data and reviewed by Van Der Maaten et al.
 \cite{Van-Der-Maaten2009Di}.
-Finally, the sparse sub-space clustering technique developed by Elhamifar and Vidal \cite{Elhamifar2013Sp} applies concepts and algorithm from the field of sparse approximation to tackle unsupervised clustering problems.
+Finally,  the sparse sub-space clustering technique developed by Elhamifar and Vidal \cite{Elhamifar2013Sp} that applies concepts and algorithm from the field of sparse approximation to tackle unsupervised clustering problems.
 
-\subsection{Main Contributions and Paper organisation}
+\subsection{Paper organisation}
 The method proposed in this paper is aimed at learning discriminative sub-spaces that allow dimensionality reduction, while at the same time enhancing the separability between classes.
-It is derived from our previous work on learning incoherent dictionaries for sparse approximation.
+It is derived from our previous work on learning incoherent dictionaries for sparse approximation \cite{Barchiesi2013Le}.
 
-Standard methods for feature transform will be briefly reviewed in Section \ref{sec:aft}, as their limitations lead to the main motivation for this work.
-To provide the context of the proposed method, the incoherent dictionary learning problem will be introduced in Section \ref{sec:idl}, while Section \ref{sec:lis} will contain the main contribution of this paper consisting in learning incoherent subspaces for classification.
+The incoherent dictionary learning problem will be introduced in Section \ref{sec:idl}, while Section \ref{sec:lis} will contain the main contribution of this paper consisting in learning incoherent subspaces for classification.
 Numerical experiments are presented in Section \ref{sec:ne}, and conclusions are drawn in Section \ref{sec:end}.
@@
 where $\NewFeas$ represents the matrix of training features after the transform stage.
-This latter approach is especially suitable when working with a large number of classes in a space of relatively small dimension (as in the numerical experiment presented in Section \ref{sec:fisher} where $\nCat=3$ and $\nDim=3$), as in this case multiple classes might be assigned to the same subspace.
+This latter approach is especially suitable when working with a large number of classes in a space of relatively small dimension, as in this case multiple classes might be assigned to the same subspace.
 \section{Numerical Experiments}\label{sec:ne}
-\subsection{Feature visualisation}
+\subsection{Feature visualisation}\label{sec:visu}
 To illustrate the \Acronym{s-ipr} algorithm for feature transform, we first run visualisation experiments depicting how different feature transform methods act on training and test data.
 \subsubsection{Synthetic data}
+\begin{figure}
+\centering
+\includegraphics[width=.6\textwidth]{./Code/Datasets/GetToyExampleDataset.pdf}
+\caption{\label{fig:toy}Synthetic data generated along one-dimensional subspaces of $\real^{2}$.}
+\end{figure}
 Figure \ref{fig:toy} displays a total of $1500$ synthetic features in $\real^{2}$ belonging to $3$ different classes that we generated for this experiment.
@@
 \begin{figure}
-\includegraphics[width=\textwidth]{./Code/Datasets/GetToyExampleDataset.pdf}
-\caption{\label{fig:toy}Synthetic data generated along one-dimensional subspaces of $\real^{2}$.}
-\end{figure}
-
-Figure
-\begin{figure}
 \includegraphics[width=\textwidth]{./Code/toyvisu.pdf}
@@
 \end{figure}
+
+Figure \ref{fig:toyNewFea} displays the result of the application of feature transforms to the data depicted in Figure \ref{fig:toy} using subspaces of dimension $1$ (with the exception of \Acronym{lda} that projects the data onto a space of dimension $\nUniCat-1=2$).
+To generate the plots, we divided the data into a training set (displayed using the `+' marker) and a test set (displayed using the `o' marker).
+Samples were drawn in random order from the dataset and assigned to either the training set or the test set, with the former containing $70\%$ of the total data and the latter containing the remaining $30\%$.
+Then, we applied feature transforms on the training set, thereby learning the transform operators, and applied them to the test set.
+
+Starting from the top-left plot, we can observe that \Acronym{PCA} identified the direction $x=y$ as the one-dimensional subspace that contains most of the variance of the training set.
+However, given the type of dataset and the dimensionality reduction caused by \Acronym{pca}, features from all classes are overlapping, making this transform a poor choice for classification.
+Similar observations can be drawn from analysing the result of \Acronym{s-pca}, although this transform identifies the direction $y=0$ as the one that leads to statistical dependence between the value of the transformed features in the training set and the relative class.
+\Acronym{lda} does not introduce any dimensionality reduction in this case, as it projects the features onto a space of dimension $\nUniCat-1=2$, leaving the original features unaltered.
+However, in the \Acronym{lda} plot we can appreciate the separation between training set and test set that is difficult to notice in the other plots.
+
+Finally, the plot at the right-bottom corner of Figure \ref{fig:toyNewFea} displays the results of the \Acronym{s-ipr} algorithm.
+In setting the parameters of \Acronym{s-ipr}, we chose a $2$ times over-complete dictionary, a number of active atoms equal to half the dimension of the data, and minimal mutual coherence.
+In the case considered here, this means $\nAto = 4$, $\nActiveAtoms = 1$ and $\coherence = \sqrt{(\nAto-\nDim)/\nDim(\nAto-1)} \approx 0.33$.
+As discussed in Section \ref{sec:iprclass}, \Acronym{s-ipr} does not project whole sets of features onto a unique sub-space, but rather learns one sub-space for each category, and projects features onto the nearest sub-space.
+The result depicted here shows that three directions were identified containing data from mostly one category each.
+Since the incoherent dictionary learning is designed to learn atoms with minimal mutual coherence, the angles between the directions of the sub-spaces learned by \Acronym{s-ipr} are approximately equal.
+Prior information regarding the directions of the data would allow to relax the parameter $\coherence$, and track more closely the directions of the three data classes.
+
 \subsubsection{Iris dataset}
 \begin{figure}
-\includegraphics[width=\textwidth]{./Code/Datasets/GetFisherIrisDataset.pdf}
-\caption{\label{fig:fisher3} First three features of the `iris' dataset depicting measurements of sepal length, sepal with and petal length of three iris species.}
+\centering
+\includegraphics[width=0.6\textwidth]{./Code/Datasets/GetFisherIrisDataset.pdf}
+\caption{\label{fig:fisher3} First three features of the `iris' dataset depicting measurements of sepal length, sepal width and petal length of three iris species.}
 \end{figure}
+Figure \ref{fig:fisher3} displays a subset of the `iris' dataset, a popular database that has been used extensively to test and benchmark classification algorithms.
+The original dataset contains measurements of the sepal length, sepal width, petal length and petal width of three species of iris, namely `setosa', `versicolor' and `virginica'.
+In this visualisation experiment we selected the first $3$ features to be able to depict the data using three dimensional scatter plots.
+From observing the distribution of the data in the feature space, we see that `setosa' is relatively separated from the other two classes, while the features relative to `virginica' and `versicolor' substantially overlap, with only a few exemplars of `virginica' being distinguishable due to large sepal length and petal length.
 \begin{figure}
-\includegraphics[width=\textwidth]{./Code/fishervisu.pdf}
-\caption{\label{fig:toyNewFea}Feature transform applied to the iris data in Figure \ref{fig:fisher3}.
+\includegraphics[width=\textwidth]{./Code/fishervisu.eps}
+\caption{\label{fig:fisher3NewFea}Feature transform applied to the iris data in Figure \ref{fig:fisher3}.
 Different colours correspond to different classes, `+' and `o' markers represent samples taken from the training and test set respectively.}
@@
 
-\subsection{Classification results}
+The results of feature transforms are depicted in Figure \ref{fig:fisher3NewFea}.
+This time, we learn $2$ dimensional subspaces from the $3$ dimensional data points and plot the transformed features, along with the learned planes.
+We observe that \Acronym{pca} identifies a direction along a diagonal axis that follows the distribution of features displayed in Figure \ref{fig:fisher3}.
+\Acronym{s-pca}
+, on the other hand, projects the features onto a horizontal plane that slightly enhances the separation between `versicolor' and `virginica' samples.
+\Acronym{lda} results in a projection where features belonging to the same category are closely clustered together, but fails to separate the classes `versicolor' and `virginica'.
+Finally, the output of \Acronym{s-ipr} displays three distinct sub-spaces associated with the three classes.
+As in the other plots, the separation between `versicolor' and `virginica' is far from perfect, however features from the  two classes are mostly projected onto the respective sub-spaces.
+Features belonging to the `setosa' category are mostly clustered together as a result of their projection onto the black subspace, however we can note a few test samples that have been associated by the algorithm to the blue sub-space.
+\subsection{Classification}\label{sec:cla}
+\begin{table}
+\centering
+\begin{tabular}{lccc}\textbf{Name} & \nDim & \nUniCat & \nFea \\
+\hline 
+Iris & 4 & 3 & 150\\
+Balance & 4 & 3 & 625\\
+Parkinsons & 23 & 2 & 197 \\
+Sonar & 60 & 2 & 208\\
+USPS & 256 & 3 & 1405
+\end{tabular}
+\caption{\label{tab:datasets}Dataset used in the classification evaluation of feature transform algorithms.
+All the datasets can be downloaded from \href{http://archive.ics.uci.edu/ml/datasets.html}{http://archive.ics.uci.edu/ml/datasets.html}.
+Note that we only use a subset of the USPS dataset containing the digits $1$, $3$ and $8$.}
+\end{table}
+
+In the previous section, we have illustrated how the \Acronym{s-ipr} algorithm is able to learn incoherent sub-spaces that model the distribution of features belonging to different classes.
+Here we evaluate \Acronym{s-ipr} and the other feature transform algorithms in the context of supervised classification.
+To perform the classification, features are transformed using the methods already used for comparison in Section \ref{sec:visu} by learning a transform operator on the training set and applying it to the test set.
+We use a $5$-fold stratified cross-validation to classify all the features in a dataset during the test stage.
+This method produces $5$ independent classification problems with a ratio between the number of training and test samples equal to $8:2$.
+Once the features have been transformed, a $k$-nearest neighbour classifier with $k=5$ is used to estimate a class.
+
+We employ the datasets detailed in Table \ref{tab:datasets}, and for each of them we evaluate the misclassification ratio, that is defined as the fraction of misclassified samples as a proportion of the total number of samples in the test set, averaged over the $5$ independent classification problems created by the stratified cross-validation protocol.
 \begin{figure}
@@
 \subfigure{\includegraphics[width=.45\textwidth]{Code/Util/sonar.pdf}}
-\caption{}
+%\subfigure{\includegraphics[width=.45\textwidth]{Code/Util/usps.pdf}}
+\caption{\label{fig:class} Classification results.}
 \end{figure}
 
-\subsection{Classifying the \Acronym{fisheriris} dataset}\label{sec:fisher}
-To illustrate the \Acronym{s-ipr} algorithm for feature transform, we apply it to the $\Acronym{fisheriris}$ dataset \cite{Fisher1936Th}.
-This contains data for the classification of $150$ iris specimens into the classes \emph{setosa}, \emph{versicolor} and \emph{virginica}.
-The features corresponding to measurements on the sepal length, sepal width and petal length for the flowers are stored in the matrix $\Feas\in\ambient^{3\times 150}$.
-As a pre-processing step, we subtract the mean and normalise the standard deviation of the features to avoid large variations or offsets in any of the dimensions of the data.
+Figure \ref{fig:class} displays for each dataset the misclassification ratio as a function of the rank of the subspace learned by the algorithms.
+In the plots `none' indicates that no feature transform was applied (hence resulting in a sub-space rank equal to the dimension of the original features).
+In general we can see that \Acronym{s-ipr} does not perform as well as the other techniques, and is only comparable at high ranks that do not achieve an overall better classification ratio.
+Starting from the `iris' dataset, \Acronym{lda} achieves the best performance followed by one-dimensional subspaces learned using \Acronym{pca}.
+Both \Acronym{s-pca} and \Acronym{s-ipr} work better when learning subspaces of high rank.
+Note that, at rank $\nDim=4$ all the methods are equivalent because they are not performing dimensionality reduction.
+The results relative to the balance dataset are similar, with again \Acronym{lda} achieving the best misclassification ratio.
+Although the results on the `Parkinsons' and `sonar' datasets present similar trends regarding \Acronym{s-ipr}, here \Acronym{lda} does not prove to be as successful as \Acronym{pca} and \Acronym{s-pca} in separating features belonging to different classes.
 
-The dictionary learning is run using the \Acronym{SMALLBox} toolbox and the Incoherent dictionary learning add-on\footnote{http://code.soundsoftware.ac.uk/projects/smallbox.}.
-We learn an over-complete dictionary $\Dic$ consisting of $\nAto=10$ atoms and a sparse approximation $\Coeff$ that uses $\nActiveAtoms=2$ active atoms for each signal.
-Incoherent sub-dictionaries are learned by the \Acronym{s-ipr} method, setting the mutual coherence to the theoretical lower bound attainable by a $\nDim\times \nAto$ dictionary $\mu_{\Label{min}} = \sqrt{(\nAto-\nDim)/\nDim(\nAto-1)}\approx 0.5$ \cite{Strohmer2003Gr}.
-A maximum number of $\nIter=20$ iterations is set as the stopping criterion.
-These parameters are within the range used in previous studies of incoherent dictionary learning \cite{Barchiesi2013Le}, and have not been optimised for this task.
+\section{Conclusion}\label{sec:end}
+\subsection{Summary}
+We have presented the \Acronym{s-ipr} algorithm for learning incoherent subspaces from data belonging to different categories.
+The encouraging experimental results obtained on the visualisation of the synthetic dataset and of a subset of features taken from the 'iris` dataset motivated us to test \Acronym{s-ipr} as a general method for feature transform to be used in classification problems.
+Unfortunately, we found that the performance of our proposed method on a group of datasets commonly used to benchmark classification methods is not competitive compared to traditional and state-of-the-art methods for feature transform.
 
-The proposed method is used to learn $\nDimSub=1$ dimensional sub-spaces, and is  compared to \Acronym{pca} with a number of components $\nComponents=2$ and to \Acronym{lda}.
-Figure \ref{fig:fisher} shows the results of the experiment.
-The original features show a clear cluster corresponding to the setosa class, and overlapping data for the other two classes.
-Applying \Acronym{pca} reduces the dimensionality of the $3$-dimensional features to a rank $2$ principal sub-space, but worsens the separation between classes.
-Both \Acronym{lda} and the proposed method are able to improve the separation of the three classes, but in remarkably different ways: while the former optimises the linear separability of different classes, features derived from \Acronym{s-ipr} are projected onto low-dimensional discriminative subspaces.
+The negative results presented in Section \ref{sec:cla} do not imply that \Acronym{s-ipr} is completely unsuitable as a tool for modelling data for classification, but they rather open a few important areas of future research that should be investigate to better understand the strengths and limitations of the proposed method.
 
-To assess the classification performance of the various techniques, we created a random $5$-fold partition of the training set used for cross-validation, and classified the transformed features using \Acronym{knn} with $5$ neighbours.
-The misclassification rate $\Acronym{mcr}$ defined as the number of misclassified samples divided by the total number of samples is displayed for the various methods in the first row of Table \ref{tab:fisher}.
-\Acronym{s-ipr} and \Acronym{lda} are both able to attain a $7\%$ misclassification error, improving on the $10\%$ obtained on the original features.
-The worst result is achieved by \Acronym{pca}, whose dimensionality reduction has the effect of mixing together features belonging to different classes, and achieves an error of $44\%$.
+\subsection{Future work}
+The main assumption made when using incoherent dictionary learning for classification is that high dimensional features are arranged onto lower-dimensional sub-spaces, and that features belonging to different classes can be modelled using different subspaces that are mutually incoherent.
+This assumption might be met by some datasets, but might not generally be satisfied by others.
+Understanding the general distribution of the features in a  dataset might be a necessary first step to inform a subsequent choice of algorithm, so that \Acronym{s-ipr} is only used in cases where its premise about the feature distribution is valid.
+This same argument holds for the whole class of linear models that comprises the dictionary learning model.
+Indeed, many feature transform techniques have equivalent kernelized versions to model non-linear data.
 
-\begin{table}[!htb]
-\begin{center}
-\begin{tabular}{|l|c|c|c|c|}
-\hline
-  & none & \Acronym{pca} & \Acronym{lda} & \Acronym{s-ipr} \\
-  \hline
-  \textbf{\Acronym{mcr} - \Acronym{fisheriris}} & 0.10 & 0.44 & 0.07 & 0.07\\
-\textbf{\Acronym{mcr} - \Acronym{usps}} & 0.343 & 0.433 & 0.359 & 0.315 \\
-\hline
-\end{tabular}
-\caption{\label{tab:fisher} Misclassification error evaluated using different feature transform methods on the \Acronym{fisheriris} and \Acronym{usps} datasets (none indicates no feature transform).}
-\end{center}
-\end{table}
-\subsection{Classification of the \Acronym{usps} digits dataset}
-In order to evaluate the proposed method on a more challenging dataset that  contains signals of higher dimension, we run a similar experiment on the \Acronym{usps} digits dataset, which consists of a collection of $16\times 16$ pixels images of hand-written digits \footnote{http://www.cs.nyu.edu/~roweis/data.html.}.
-We selected the digits \Acronym{1}, \Acronym{3} and \Acronym{8} resulting in a total of $1405$ examples stored in the matrix $\Feas\in\ambient^{256\times 1405}$.
-
-We investigated the effect of the parameters \\ $\nAto=\curlyb{256,512,1024}$, $\nDimSub=\curlyb{2,4,8,16,32,64,128}$ and $\nActiveAtoms = \curlyb{2,25,128,256}$ on the misclassification error, and compared the \Acronym{s-ipr} to \Acronym{pca} and \Acronym{lda}.
-For \Acronym{pca}, the number of principal components $\nComponents$ is automatically set by the algorithm to choose an approximation that retains $95\%$ of the variance of the data.
-The second row of Table \ref{tab:fisher} shows the results obtained with the best choice of parameters $\optimal{\nAto}=512$ (corresponding to a 2 times over-complete dictionary), $\optimal{\nDimSub}=64$ and $\optimal{\nActiveAtoms}=128$.
-In this case, \Acronym{s-ipr} achieves a $31.5\%$ misclassification error, while the other two techniques do not improve the \Acronym{k-nn} classification on to the original features, which scores $34.3\%$.
-
-Analysing the \Acronym{mcr} as a function of the parameters $\nAto$, $\nDimSub$ and $\nActiveAtoms$, the most relevant trend observed empirically on this dataset is that a value $\nDimSub<16$ steeply increases the misclassification error.
-This suggests that the dimensionality of the incoherent sub-spaces must be large enough to retain discriminative characteristics of the original high-dimensional data.
-The dependency of \Acronym{mcr} on the other two parameters is less significant, suggesting that dictionary learning leads to discriminative subspaces regardless of the number of active atoms $\nActiveAtoms$ or the number of atoms in the dictionary $\nAto$.
-\section{Conclusion}\label{sec:end}
-We have presented the \Acronym{s-ipr} algorithm for learning incoherent subspaces, and employed it as a feature transform method in the context of supervised classification.
-The encouraging experimental results obtained on the \Acronym{fisheriris} and the \Acronym{usps} datasets suggest that the proposed algorithm can overcome limitations of standard feature transforms methods, making it a viable tool for machine learning research and practice.
-A better theoretical understanding of the performance of \Acronym{s-ipr} (especially in relation to other methods for subspace and manifold learning), and further validations of the technique on challenging classification problems constitute the most promising avenues for future research.
+Other substantial improvements can be made on the algorithm itself.
+The present implementation of \Acronym{s-ipr} contains a fixed parameter $\coherence$ that promotes minimal mutual coherence between the sub-spaces used to approximate different data classes.
+Knowledge about the distribution of the features might lead to relaxing this parameter, learning sub-spaces that are closer to the true distribution of the features and in turn improving class separation.
+Moreover, different values of mutual coherence for different pairs of subspaces can be easily included in the optimisation, greatly enhancing the flexibility of \Acronym{s-ipr} as a modelling tool.
 

### 2014-01-15 code/TestClassSubSpaRan.m

@@ -29,7 +29,7 @@ for iDat=1:length(datasets)
     if ~isfloat(fea), fea = double(fea); end    %transform features to float if needed
     fprintf('done\n');
     feaDim = size(fea,2);
-    subSpaRanks = rmd(ceil(linspace(1,feaDim,20)));
+    subSpaRanks = rmd(ceil(linspace(1,feaDim,22)));
     for iSSR = 1:length(subSpaRanks)
         fprintf('testing transforms using subspaces of rank %d... ',subSpaRanks(iSSR));
         par.subSpaRan = subSpaRanks(iSSR);

### 2014-01-15 code/TestTransforms.m

@@ -4,10 +4,10 @@ addpath(genpath(pwd)); %add path to matlab search path
 
 %% Parameters and defaults
 if ~exist('fea','var') || isempty(fea) || ~exist('cat','var') || isempty(cat)
-    [fea,cat] = GetToyExampleDataset;
+    [fea,cat] = GetFisherIrisDataset;
     if isnumeric(cat), cat = cellstr(num2str(cat)); end
-%     nDim = 3; %data has 4 dimensions
-%     fea = fea(:,1:nDim);
+    nDim = 3; %data has 4 dimensions
+    fea = fea(:,1:nDim);
 end
 
 if ~exist('par','var') || isempty(par), par = struct; end
@@ -15,9 +15,9 @@ if ~exist('par','var') || isempty(par), par = struct; end
 options = statset('UseParallel',false); %set option to use parallel computing
 
 def.nAto = 2*size(fea,2); %number of atoms (twice overcomplete dictionary)
-def.subSpaRan = 1; %subspace range
+def.subSpaRan = 2; %subspace range
 def.perActAto = 50; %percentage of active atoms
-def.nKNNs = 5; %number of nearest neighbours for classification
+def.nKNNs = 1; %number of nearest neighbours for classification
 def.visu = true;
 
 par = setdefaultoptions(par,def);
@@ -178,13 +178,19 @@ for iCat=1:length(uniCat) %for every category
     ind = strcmp(uniCat(iCat),traCat); %find indexes of data in the trainig set belonging to iCat category
     scatter3(traFea(ind,1),traFea(ind,2),traFea(ind,3),size,col{iCat},'Marker','+'); %scatter data
     hold on
-    if strcmp(method,'IPR')
-        scale = 3;
-        quiver3(0,0,0,scale*subs{iCat}(1),scale*subs{iCat}(2),scale*subs{iCat}(3));
-        quiver3(0,0,0,-scale*subs{iCat}(1),-scale*subs{iCat}(2),-scale*subs{iCat}(3));
-    end
+%     if strcmp(method,'IPR')
+%         scale = 3;
+%         quiver3(0,0,0,scale*subs{iCat}(1),scale*subs{iCat}(2),scale*subs{iCat}(3));
+%         quiver3(0,0,0,-scale*subs{iCat}(1),-scale*subs{iCat}(2),-scale*subs{iCat}(3));
+%     end
[…]

### 2014-01-15 code/Util/drawplane.m

@@ -1,0 +1,31 @@
+function drawplane(Y,c)
+if ~nargin, unittest; return; end
+if nargin<2, c='k'; end
+
+a = Y(:,1:3); % select 3 points in Y
+% find normal vector
+a2ma1 = a(:,2) - a(:,1);
+a3ma1 = a(:,3) - a(:,1);
+n = cross(a2ma1,a3ma1);
+% plot the plane
+d = n'*a(:,1);
+alpha = 1.5;
+xmin = alpha*min(Y(1,:));
+xmax = alpha*max(Y(1,:));
+ymin = alpha*min(Y(2,:));
+ymax = alpha*max(Y(2,:));
+x = linspace(xmin,xmax,10);
+y = linspace(ymin,ymax,10);
+[X,Y] = meshgrid(x,y);
+Z = (d - (n(1)*X) - (n(2)*Y))/n(3);
+h = surf(X,Y,Z);
+set(h,'FaceColor',c,'FaceAlpha',0.3,'LineStyle','none');
+
+function unittest
+close all
+X = randn(3,100);
+D = pca(X');
+Y = D(:,1:2)*pinv(D(:,1:2))*X;
+scatter3(Y(1,:),Y(2,:),Y(3,:));
+hold on
+drawplane(Y);

### 2014-01-15 manuscript.tex
improved mcr plot

@@
 In this article we present the supervised iterative projections and rotations (\Acronym{s-ipr}) algorithm, a method for learning discriminative incoherent subspaces from data.
-We derive \Acronym{s-ipr} as a supervised extension of our previously developed iterative projections and rotations (\Acronym{ipr}) algorithm for incoherent dictionary learning, and we employ it to learn incoherent sub-spaces that model signals belonging to different classes.
+We derive \Acronym{s-ipr} as a supervised extension of our previously proposed iterative projections and rotations (\Acronym{ipr}) algorithm for incoherent dictionary learning, and we employ it to learn incoherent sub-spaces that model signals belonging to different classes.
 We test our method as a feature transform for supervised classification, first by visualising transformed features from a synthetic dataset and from the `iris' dataset, then by using the resulting features in a classification experiment.
@@
 The union of incoherent sub-spaces model proposed by Schnass and Vandergheynst \cite{Schnass2010A-} employes a very similar intuition to the one that inspired our proposed method, and models features belonging to different classes using incoherent subspaces.
-
 Other methods for supervised dimensionality reduction include metric learning algorithms \cite{xing2002distance}, sufficient dimensionality reduction \cite{li1991sliced} and Bair's supervised principal components \cite{Bair06predictionby}.
@@
 \centering
-\subfigure{\includegraphics[width=.45\textwidth]{Code/Util/fisheriris.pdf}}
-\subfigure{\includegraphics[width=.45\textwidth]{Code/Util/balance.pdf}}
-\subfigure{\includegraphics[width=.45\textwidth]{Code/Util/parkinsons.pdf}}
-\subfigure{\includegraphics[width=.45\textwidth]{Code/Util/sonar.pdf}}
-%\subfigure{\includegraphics[width=.45\textwidth]{Code/Util/usps.pdf}}
-\caption{\label{fig:class} Classification results.}
+\subfigure{\includegraphics[width=.48\textwidth]{Code/Util/fisheriris.pdf}}
+\subfigure{\includegraphics[width=.48\textwidth]{Code/Util/balance.pdf}}
+\subfigure{\includegraphics[width=.48\textwidth]{Code/Util/parkinsons.pdf}}
+\subfigure{\includegraphics[width=.48\textwidth]{Code/Util/sonar.pdf}}
+%\subfigure{\includegraphics[width=.7\textwidth]{Code/Util/usps.pdf}}
+\caption{\label{fig:class} Misclassification ratio as a function of the rank of the subspace employed during feature transforms for the datasets  `iris',`balance',`Parkinsons' and `sonar'}
 \end{figure}
@@
 The encouraging experimental results obtained on the visualisation of the synthetic dataset and of a subset of features taken from the 'iris` dataset motivated us to test \Acronym{s-ipr} as a general method for feature transform to be used in classification problems.
-Unfortunately, we found that the performance of our proposed method on a group of datasets commonly used to benchmark classification methods is not competitive compared to traditional and state-of-the-art methods for feature transform.
+Unfortunately, we found that the performance of our proposed method on a group of datasets commonly used to benchmark classification algorithms is only competitive compared to traditional and state-of-the-art methods for feature transform at high sub-space ranks.
 
@@
 This assumption might be met by some datasets, but might not generally be satisfied by others.
-Understanding the general distribution of the features in a  dataset might be a necessary first step to inform a subsequent choice of algorithm, so that \Acronym{s-ipr} is only used in cases where its premise about the feature distribution is valid.
+Understanding the general distribution of the features in a  dataset might be a necessary first step to inform a subsequent choice of algorithm, so that \Acronym{s-ipr} can be used in cases where its premise about the feature distribution is valid.
 This same argument holds for the whole class of linear models that comprises the dictionary learning model.
-Indeed, many feature transform techniques have equivalent kernelized versions to model non-linear data.
+Indeed, many feature transform techniques have equivalent \emph{kernelized} versions to model non-linear data.
 

### 2014-01-15 code/TestClassSubSpaRan.m

@@ -37,6 +37,6 @@ for iDat=1:length(datasets)
         fprintf('done\n');
     end
     mcrs = [res(iDat,:).mcr]';
-    figure, PlotMCRs(mcrs,subSpaRanks,datasets{iDat});
+    figure, PlotMCRs(mcrs,subSpaRanks,length(unique(cat)),datasets{iDat});
 end
 

### 2014-01-15 code/Util/PlotMCRs.m

@@ -1,13 +1,17 @@
-function PlotMCRs(mcrs,subSpaRanks,dataset)
+function PlotMCRs(mcrs,subSpaRanks,nUniCat,dataset)
 methodNames = {'none','LDA','PCA','S-PCA','S-IPR'};
 plot(subSpaRanks(end),mean(mcrs(:,1)),'ks','MarkerSize',10)
 hold on
-plot(subSpaRanks(end)-1,mean(mcrs(:,4)),'rs','MarkerSize',10)
+plot(nUniCat-1,mean(mcrs(:,4)),'rs','MarkerSize',10)
 plot(subSpaRanks,mcrs(:,[1,2,5]),'--o','MarkerSize',10)
 legend(methodNames);
 xlabel('Sub-space rank');
 ylabel('misclassification ratio');
 title(sprintf('Classification of %s dataset',dataset));
+randomChoice = 1/nUniCat;
+ylim([0,max([randomChoice, max(mcrs(:))])*1.2]);
+plot(get(gca,'XLim'),[randomChoice,randomChoice],'k-');
+text(min(get(gca,'xlim')),randomChoice*1.1,'chance level','Margin',10);
 set(gca,'YGrid','on','XGrid','off',...
     'XTick',rmd([subSpaRanks(1:end-1), subSpaRanks(end)-1, subSpaRanks(end)]));
 a = fileparts(which(mfilename));
